---
title: "4. COV002 CoP analysis - non-parametric"
output: html_document
---

*Summary*
This file provides the code for running the non-parametric Rstan model for the COV002 data. Separate models are run for:

- Neutralising antibodies
- Spike binding antibodies
- RBD binding antibodies



*Notes*
The final models are generated with joint_analysis_CoP_nonparametric_jan24_priors_oxford.stan, (priors = a_raw ~ normal(0, 1); a0 ~ normal(-5, 2)). 

- fit_neut_nonpara2_efficacy_cov002_CoP_march_3knots_log10.rds
- fit_spike_nonpara2_efficacy_cov002_CoP_march_3knots_log10.rds
- fit_rbd_nonpara2_efficacy_cov002_CoP_march_3knots_log10.rds

Non-weighted sensitivity:

- fit_neut_nonpara2_efficacy_cov002_CoP_march_3knots_log10_noweight.rds
- fit_spike_nonpara2_efficacy_cov002_CoP_march_3knots_log10_noweight.rds
- fit_rbd_nonpara2_efficacy_cov002_CoP_march_3knots_log10_noweight.rds

# 1. Setup

```{r setup 1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#desktop
knitr::opts_knit$set(root.dir = "C:/Users/lucyr/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#laptop
```

```{r setup 2}
#setwd("C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#desktop
setwd("C:/Users/lucyr/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#laptop
getwd()
library("ggplot2")
library("rstan")
library("gridExtra")
library("bayesplot")
library("tidyverse")
library("loo")
library("bridgesampling")
library("ggridges")
library("data.table")
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

```{r define functions}
lci<-function(x){quantile(x, 0.025)}
uci<-function(x){quantile(x, 0.975)}
```

```{r complete case subsets}
#Neutralising antibodies
neut_cc_data<-model_data%>%
  subset(!is.na(neuts.PB28_log))
num_subjects_neuts<-nrow(neut_cc_data)
table(neut_cc_data$outcome, neut_cc_data$vaccinated)

#Spike IgG
spike_cc_data<-model_data%>%
  subset(!is.na(igg.S.PB28_log))
num_subjects_spike<-nrow(spike_cc_data)
table(spike_cc_data$outcome, spike_cc_data$vaccinated)

#RBD IgG
RBD_cc_data<-model_data%>%
  subset(!is.na(igg.RBD.PB28_log))
num_subjects_RBD<-nrow(RBD_cc_data)
table(RBD_cc_data$outcome, RBD_cc_data$vaccinated)
```

# 2. Models

## a. Neutralising antibodies

The data is structured with a row for each trial participant, and columns of their socio-demographic and clinical characteristics, as well as their trial measurements and outcomes.

The covariates included are:
- HCW_0 = Healthcare worker seeing <1 COVID-19 patient per day on average (1=yes, 0=no)
- HCW_1plus = Healthcare worker seeing 1 or more COVID-19 patient per day on average (1=yes, 0=no)
- sc_age_cen = age (years), centred on the mean age
- obese = BMI 30+  (1=yes, 0=no)
- white = white ethnicity  (1=yes, 0=no)
- comorbidity = comorbidity (1=yes, 0=no)

The weights are defined in the neut_weight variable. For sensitivity analysis in which the model was not weighted, all weights were set to 1.

```{r data setup neuts}
neuts<-c(seq(min(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1]), round(max(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1]), 1)+0.0103,0.1)) #extended the sequence  because it was cutting it slightly short
Z_range<-seq(min(neuts),max(neuts),0.1)
dX<-0.1
Z_pos<-floor((neut_cc_data$neuts.PB28_log10-Z_range[1])/dX)
Z_pos[Z_pos<1]<-1
Z_pos[Z_pos>length(Z_range)]<-length(Z_range)

#senstivity without weights
#neut_cc_data$neut_weight<-1
```


```{r model run neuts}
trial_dat <- list(N=num_subjects_neuts, 
                  C=neut_cc_data$C,
                  A=neut_cc_data$A,
                  vaccinated=neut_cc_data$vaccinated,
                  Z=neut_cc_data$neuts.PB28_log10,
                  pers_yrs_at_risk=neut_cc_data$pers_yrs_at_risk_cop,
                  num_knots=3,
                  knots=unname(quantile(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1],probs=seq(from=0, to=1, length.out = 3))), 
                  spline_degree=3,
                  X=Z_range, 
                  spline_length=length(Z_range),
                  dX=dX,
                  Z_index=Z_pos,
                  weight=neut_cc_data$neut_weight,
                  cov1=neut_cc_data$HCW_0,
                  cov2=neut_cc_data$HCW_1plus,
                  cov3=neut_cc_data$sc_age_cen,
                  cov4=neut_cc_data$obese,
                  cov5=neut_cc_data$white,
                  cov6=neut_cc_data$comorbidity
                  
)
setwd("C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#desktop
getwd()
model <- stan_model(file = 'Nonparametric_model_cov002.stan')
fit_neut_nonpara <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log10_lik(fit_neut_nonpara)
loo <- loo(log_lik)
print(loo)
#visualise parameters
pars = c("gam",  "alpha", "cVE_hat[1]", "cVE_pr_hat[1]", "cVE_hat[10]", "cVE_pr_hat[10]", "cVE_hat[100]", "cVE_pr_hat[100]")
pars = c("gam",  "alpha", "cVE_hat[1]", "cVE_pr_hat[1]", "a0", "a0_pr", "tau", "tau_pr", "a_raw", "a_raw_pr")
print(fit_neut_nonpara, pars=pars)
traceplot(fit_neut_nonpara, pars=pars, inc_warmup = TRUE)
pairs(fit_neut_nonpara, pars=pars)
```

```{r save as RDS}
fit_neut_nonpara@stanmodel@dso <- new("cxxdso")
saveRDS(fit_neut_nonpara, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_neut_nonpara2_efficacy_cov002_CoP_march_3knots_log10_noweight.rds")
fit<-fit_neut_nonpara
```


## b. Spike antibodies

The same models as above are run for the Spike binding antibodies outcome.

```{r data setup spike}
spikes<-seq(min(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1]), round(max(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1])+0.147761, 1), 0.1)
Z_range<-seq(min(spikes),max(spikes),0.1)
dX<-0.1
Z_pos<-floor((spike_cc_data$igg.S.PB28_log10-Z_range[1])/dX)
Z_pos[Z_pos<1]<-1
Z_pos[Z_pos>length(Z_range)]<-length(Z_range)
#senstivity without weights
#spike_cc_data$spike_rbd_weight<-1
```


```{r model run spike}
trial_dat <- list(N=num_subjects_spike, 
                  C=spike_cc_data$C,
                  A=spike_cc_data$A,
                  vaccinated=spike_cc_data$vaccinated,
                  Z=spike_cc_data$igg.S.PB28_log10,
                  pers_yrs_at_risk=spike_cc_data$pers_yrs_at_risk_cop,
                  num_knots=3,
                  knots=unname(quantile(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1],probs=seq(from=0, to=1, length.out = 3))), 
                  spline_degree=3,
                  X=Z_range, 
                  spline_length=length(Z_range),
                  dX=dX,
                  Z_index=Z_pos,
                  weight=spike_cc_data$spike_rbd_weight,
                  cov1=spike_cc_data$HCW_0,
                  cov2=spike_cc_data$HCW_1plus,
                  cov3=spike_cc_data$sc_age_cen,
                  cov4=spike_cc_data$obese,
                  cov5=spike_cc_data$white,
                  cov6=spike_cc_data$comorbidity
                  
)
setwd("C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#desktop
getwd()
model <- stan_model(file = 'Nonparametric_model_cov002.stan')
fit_spike_nonpara <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r}
#log likelihood
log_lik <- extract_log10_lik(fit_spike_nonpara)
loo <- loo(log_lik)
print(loo)
#visualise parameters
pars = c("gam",  "alpha", "cVE_hat[1]", "cVE_pr_hat[1]", "cVE_hat[10]", "cVE_pr_hat[10]", "cVE_hat[100]", "cVE_pr_hat[100]")
pars = c("gam",  "alpha_ref", "cVE_hat[1]", "cVE_pr_hat[1]", "a0", "a0_pr", "a_raw", "a_raw_pr")
pars = c("gam",  "alpha_ref")
print(fit_spike_nonpara, pars=pars)
traceplot(fit_spike_nonpara, pars=pars, inc_warmup = TRUE)
pairs(fit_spike_nonpara, pars=pars)
```


```{r save as RDS}
fit_spike_nonpara@stanmodel@dso <- new("cxxdso")
saveRDS(fit_spike_nonpara, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_spike_nonpara_cov002.rds")
```

## c. RBD antibodies

The same models as above are run for the RBD binding antibodies outcome.

```{r data setup RBD}
RBDs<-seq(min(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1]), round(max(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1]), 1)+0.1086,0.1)
Z_range<-seq(min(RBDs),max(RBDs),0.1)
dX<-0.1
Z_pos<-floor((RBD_cc_data$igg.RBD.PB28_log10-Z_range[1])/dX)
Z_pos[Z_pos<1]<-1
Z_pos[Z_pos>length(Z_range)]<-length(Z_range)
#sensitivity no weights
#RBD_cc_data$spike_rbd_weight<-1
```


```{r baseline model run RBD}
trial_dat <- list(N=num_subjects_RBD, 
                  C=RBD_cc_data$C,
                  A=RBD_cc_data$A,
                  vaccinated=RBD_cc_data$vaccinated,
                  Z=RBD_cc_data$igg.RBD.PB28_log10,
                  pers_yrs_at_risk=RBD_cc_data$pers_yrs_at_risk_cop,
                  num_knots=5,
                  knots=unname(quantile(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1],probs=seq(from=0, to=1, length.out = 5))), 
                  spline_degree=3,
                  X=Z_range, 
                  spline_length=length(Z_range),
                  dX=dX,
                  Z_index=Z_pos,
                  weight=RBD_cc_data$spike_rbd_weight,
                  cov1=RBD_cc_data$HCW_0,
                  cov2=RBD_cc_data$HCW_1plus,
                  cov3=RBD_cc_data$sc_age_cen,
                  cov4=RBD_cc_data$obese,
                  cov5=RBD_cc_data$white,
                  cov6=RBD_cc_data$comorbidity
                  
)
setwd("C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#desktop
getwd()
model <- stan_model(file = 'Nonparametric_model_cov002.stan')
fit_RBD_nonpara <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log10_lik(fit_RBD_nonpara)
loo <- loo(log_lik)
print(loo)
#visualise parameters
pars = c("gam",  "alpha", "cVE_hat[1]", "cVE_pr_hat[1]", "cVE_hat[10]", "cVE_pr_hat[10]", "cVE_hat[100]", "cVE_pr_hat[100]")
pars = c("gam",  "alpha", "cVE_hat[1]", "cVE_pr_hat[1]", "a0", "a0_pr", "a_raw", "a_raw_pr")
pars = c("gam",  "alpha_ref")
print(fit_RBD_nonpara, pars=pars)
traceplot(fit_RBD_nonpara, pars=pars, inc_warmup = TRUE)
pairs(fit_RBD_nonpara, pars=pars)
```


```{r save as RDS}
fit_RBD_nonpara@stanmodel@dso <- new("cxxdso")
saveRDS(fit_RBD_nonpara, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_RBD_nonpara_cov002.rds")
```

