---
title: "COV002 CoP analysis - parametric"
output: html_document
author: "Lucy Williams"
---

*Summary*
This file provides the code for running the parametric Rstan model for the COV002 data. Separate models are run for:

- Neutralising antibodies
- Spike binding antibodies
- RBD binding antibodies

The baseline model for each was the one used to generate the main results in the manuscript. 


We also ran models in which we set the slope parameter, k, to 0 for either VEin or VEpr. This was for the Savage-Dickey ratio calculation. The densities of the prior and posterior distributions at k = 0 were used to calculate the Savage-Dickey density ratio (see supplementary methods).


# 1. Setup

```{r setup}
library("ggplot2")
library("rstan")
library("gridExtra")
library("bayesplot")
library("tidyverse")
library("loo")
library("bridgesampling")
library("ggridges")
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

```{r define functions}
lci<-function(x){quantile(x, 0.025)}
uci<-function(x){quantile(x, 0.975)}
```


```{r complete case subsets}
#Neutralising antibodies
neut_cc_data<-model_data%>%
  subset(!is.na(neuts.PB28_log))
num_subjects_neuts<-nrow(neut_cc_data)
table(neut_cc_data$outcome, neut_cc_data$vaccinated)

#Spike IgG
spike_cc_data<-model_data%>%
  subset(!is.na(igg.S.PB28_log))
num_subjects_spike<-nrow(spike_cc_data)
table(spike_cc_data$outcome, spike_cc_data$vaccinated)

#RBD IgG
RBD_cc_data<-model_data%>%
  subset(!is.na(igg.RBD.PB28_log))
num_subjects_RBD<-nrow(RBD_cc_data)
table(RBD_cc_data$outcome, RBD_cc_data$vaccinated)
```


# 2. Models

## a. Neutralising antibodies

### Baseline model

The data is structured with a row for each trial participant, and columns of their socio-demographic and clinical characteristics, as well as their trial measurements and outcomes.

The covariates included are:
- HCW_0 = Healthcare worker seeing <1 COVID-19 patient per day on average (1=yes, 0=no)
- HCW_1plus = Healthcare worker seeing 1 or more COVID-19 patient per day on average (1=yes, 0=no)
- sc_age_cen = age (years), centred on the mean age
- obese = BMI 30+  (1=yes, 0=no)
- white = white ethnicity  (1=yes, 0=no)
- comorbidity = comorbidity (1=yes, 0=no)

The weights are defined in the neut_weight variable. For sensitivity analysis in which the model was not weighted, all weights were set to 1.

```{r baseline model run}
#neut_cc_data$neut_weight<-1 #for sensitivity analysis, reset the weights to 1 and save as ..._noweights
#Define data
trial_dat <- list(N=num_subjects_neuts,                               #N subjects in analysis population with neutralising antibody measurement
                  C=neut_cc_data$C,                                   #N symptomatic infections 
                  A=neut_cc_data$A,                                   #N asymptomatic infections 
                  vaccinated=neut_cc_data$vaccinated,                 #Vaccinated indicator variable (1=vaccine, 0=placebo) 
                  Z=neut_cc_data$neuts.PB28_log10,                    #Neutralising antibody titre, log10 transformed
                  pers_yrs_at_risk=neut_cc_data$pers_yrs_at_risk_cop, #Person-years at risk for CoP analysis
                  min_Z=min(neut_cc_data$neuts.PB28_log10),           #Minimum transformed CoP titre
                  max_Z=max(neut_cc_data$neuts.PB28_log10),           #Maximum transformed CoP titre
                  weight=neut_cc_data$neut_weight,                    #Inverse probability weight (1/probability of participant having a neutralising antibody titre)
                  cov1=neut_cc_data$HCW_0,                            #Healthcare worker seeing <1 COVID-19 patient per day on average (1=yes, 0=no)
                  cov2=neut_cc_data$HCW_1plus,                        #Healthcare worker seeing 1 or more COVID-19 patient per day on average (1=yes, 0=no)
                  cov3=neut_cc_data$sc_age_cen,                       #Age (years), centred on the mean age
                  cov4=neut_cc_data$obese,                            #BMI 30+  (1=yes, 0=no)
                  cov5=neut_cc_data$white,                            #white ethnicity  (1=yes, 0=no)
                  cov6=neut_cc_data$comorbidity                       #Comorbidity (1=yes, 0=no)
)

model <- stan_model(file = 'Parametric_model_cov002.stan')
fit_neut_para <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#Log likelihood
log_lik <- extract_log_lik(fit_neut_para)
loo <- loo(log_lik)
print(loo)

#View parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_in", "kk_pr", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_neut_para, pars=pars)
traceplot(fit_neut_para, pars=pars, inc_warmup = TRUE)
pairs(fit_neut_para, pars=pars)
pairs(fit_neut_para, pars=c("VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean" ))
pairs(fit_neut_para, pars=c("VE_direct_in", "VE_direct_pr", "p_direct_in", "p_direct_pr" ))
```

```{r save as RDS}
fit_neut_para@stanmodel@dso <- new("cxxdso")
#save baseline model
#saveRDS(fit_neut_para, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_neut_para_cov002.rds")
#save no weights model
#saveRDS(fit_neut_para, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_neut_para_cov002_noweights.rds")
```

```{r posterior distributions}
est_params<-data.frame(rstan::extract(fit_neut_para, pars=pars))
hist(est_params$kk_in, br=100)
hist(est_params$kk_pr, br=100)
hist(est_params$kk_in, br=1000, xlim = c(0,3))
hist(est_params$kk_pr, br=1000, xlim = c(0,3))
hist(est_params$n_50_in, br=100)
hist(est_params$n_50_pr, br=100)
hist(est_params$VE_direct_in, br=100)
hist(est_params$VE_direct_pr, br=100)
hist(est_params$VE_indirect_in, br=100)
hist(est_params$VE_indirect_pr, br=100)
hist(est_params$ps, br=100)
hist(est_params$scale_in, br=100)
hist(est_params$scale_pr, br=100)
hist(est_params$p_direct_in, br=100)
hist(est_params$p_direct_pr, br=100)
```

```{r summarise}
est_params_summary<-est_params %>%
  summarise(
    mean_ps=mean(ps),
    mean_foi_ref=mean(foi_ref),
    mean_foi_cov1=mean(foi_cov1),
    mean_foi_cov2=mean(foi_cov2),
    mean_foi_cov3=mean(foi_cov3),
    mean_foi_cov4=mean(foi_cov4),
    mean_foi_cov5=mean(foi_cov5),
    mean_foi_cov6=mean(foi_cov6),
    mean_kk_in=mean(kk_in),
    mean_kk_pr=mean(kk_pr),
    mean_n_50_in=mean(n_50_in),
    mean_n_50_pr=mean(n_50_pr),
    mean_scale_in=mean(scale_in),
    mean_scale_pr=mean(scale_pr),
    mean_VE_direct_in=mean(VE_direct_in),
    mean_VE_direct_pr=mean(VE_direct_pr),
    mean_VE_indirect_in=mean(VE_indirect_in_mean),
    mean_VE_indirect_pr=mean(VE_indirect_pr_mean),
    mean_p_direct_in=mean(p_direct_in),
    mean_p_direct_pr=mean(p_direct_pr),
    mean_VE_in=mean(VE_in_mean),
    mean_VE_pr=mean(VE_pr_mean),
    mean_VE_sym=mean(VE_sym_mean),
    mean_VE_asym=mean(VE_asym_mean),
    
    median_ps=median(ps),
    median_foi_ref=median(foi_ref),
    median_foi_cov1=median(foi_cov1),
    median_foi_cov2=median(foi_cov2),
    median_foi_cov3=median(foi_cov3),
    median_foi_cov4=median(foi_cov4),
    median_foi_cov5=median(foi_cov5),
    median_foi_cov6=median(foi_cov6), 
    median_kk_in=median(kk_in),
    median_kk_pr=median(kk_pr),
    median_n_50_in=median(n_50_in),
    median_n_50_pr=median(n_50_pr),
    median_scale_in=median(scale_in),
    median_scale_pr=median(scale_pr),
    median_VE_direct_in=median(VE_direct_in),
    median_VE_direct_pr=median(VE_direct_pr),
    median_VE_indirect_in=median(VE_indirect_in_mean),
    median_VE_indirect_pr=median(VE_indirect_pr_mean),
    median_p_direct_in=median(p_direct_in),
    median_p_direct_pr=median(p_direct_pr),
    median_VE_in=median(VE_in_mean),
    median_VE_pr=median(VE_pr_mean),
    median_VE_sym=median(VE_sym_mean),
    median_VE_asym=median(VE_asym_mean),
    
    lci_ps=quantile(ps, 0.025),
    lci_foi_ref=quantile(foi_ref, 0.025),
    lci_foi_cov1=quantile(foi_cov1, 0.025),
    lci_foi_cov2=quantile(foi_cov2, 0.025),
    lci_foi_cov3=quantile(foi_cov3, 0.025),
    lci_foi_cov4=quantile(foi_cov4, 0.025),
    lci_foi_cov5=quantile(foi_cov5, 0.025),
    lci_foi_cov6=quantile(foi_cov6, 0.025),
    lci_kk_in=quantile(kk_in, 0.025),
    lci_kk_pr=quantile(kk_pr, 0.025),
    lci_n_50_in=quantile(n_50_in, 0.025),
    lci_n_50_pr=quantile(n_50_pr, 0.025),
    lci_scale_in=quantile(scale_in, 0.025),
    lci_scale_pr=quantile(scale_pr, 0.025),
    lci_VE_direct_in=quantile(VE_direct_in, 0.025),
    lci_VE_direct_pr=quantile(VE_direct_pr, 0.025),
    lci_VE_indirect_in=quantile(VE_indirect_in_mean, 0.025),
    lci_VE_indirect_pr=quantile(VE_indirect_pr_mean, 0.025),
    lci_p_direct_in=quantile(p_direct_in, 0.025),
    lci_p_direct_pr=quantile(p_direct_pr, 0.025),
    lci_VE_in=quantile(VE_in_mean, 0.025),
    lci_VE_pr=quantile(VE_pr_mean, 0.025),
    lci_VE_sym=quantile(VE_sym_mean, 0.025),
    lci_VE_asym=quantile(VE_asym_mean, 0.025),
    
        
    uci_ps=quantile(ps, 0.975),
    uci_foi_ref=quantile(foi_ref, 0.975),
    uci_foi_cov1=quantile(foi_cov1, 0.975),
    uci_foi_cov2=quantile(foi_cov2, 0.975),
    uci_foi_cov3=quantile(foi_cov3, 0.975),
    uci_foi_cov4=quantile(foi_cov4, 0.975),
    uci_foi_cov5=quantile(foi_cov5, 0.975),
    uci_foi_cov6=quantile(foi_cov6, 0.975),
    uci_kk_in=quantile(kk_in, 0.975),
    uci_kk_pr=quantile(kk_pr, 0.975),
    uci_n_50_in=quantile(n_50_in, 0.975),
    uci_n_50_pr=quantile(n_50_pr, 0.975),
    uci_scale_in=quantile(scale_in, 0.975),
    uci_scale_pr=quantile(scale_pr, 0.975),
    uci_VE_direct_in=quantile(VE_direct_in, 0.975),
    uci_VE_direct_pr=quantile(VE_direct_pr, 0.975),
    uci_VE_indirect_in=quantile(VE_indirect_in_mean, 0.975),
    uci_VE_indirect_pr=quantile(VE_indirect_pr_mean, 0.975),
    uci_p_direct_in=quantile(p_direct_in, 0.975),
    uci_p_direct_pr=quantile(p_direct_pr, 0.975),
    uci_VE_in=quantile(VE_in_mean, 0.975),
    uci_VE_pr=quantile(VE_pr_mean, 0.975),
    uci_VE_sym=quantile(VE_sym_mean, 0.975),
    uci_VE_asym=quantile(VE_asym_mean, 0.975)
  )%>%
  round(digits=2)
est_params_summary
```


```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_in))+
  geom_histogram(binwidth = 0.02, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_in))+
  geom_histogram(binwidth = 0.2, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian()

ggplot(est_params, aes(x=kk_pr))+
  geom_histogram(binwidth = 0.02, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_pr))+
  geom_histogram(binwidth = 0.2, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian()
```

```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_in))+
  geom_density()+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_in))+
  geom_density()+
  theme_bw()+coord_cartesian()
```

```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_pr))+
  geom_density()+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_pr))+
  geom_density()+
  theme_bw()+coord_cartesian()
```


```{r read model}
fit_neut_para<-readRDS("fit_neut_para_cov002.rds")
```

### kkin 0 model

Here the same model is run, except the slope parameter, k, for VEin is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEin.

```{r kkin0 model run}
trial_dat <- list(N=num_subjects_neuts, 
                  C=neut_cc_data$C,
                  A=neut_cc_data$A,
                  vaccinated=neut_cc_data$vaccinated,
                  Z=neut_cc_data$neuts.PB28_log10,
                  pers_yrs_at_risk=neut_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(neut_cc_data$neuts.PB28_log10),
                  max_Z=max(neut_cc_data$neuts.PB28_log10),
                  weight=neut_cc_data$neut_weight,
                  cov1=neut_cc_data$HCW_0,
                  cov2=neut_cc_data$HCW_1plus,
                  cov3=neut_cc_data$sc_age_cen,
                  cov4=neut_cc_data$obese,
                  cov5=neut_cc_data$white,
                  cov6=neut_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kkin0.stan')

fit_neut_para_kkin0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_neut_para_kkin0)
loo <- loo(log_lik)
print(loo)

#visualise paramaters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_pr", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_neut_para_kkin0, pars=pars)
traceplot(fit_neut_para_kkin0, pars=pars, inc_warmup = TRUE)
pairs(fit_neut_para_kkin0, pars=pars)
```


```{r save as RDS}
fit_neut_para_kkin0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_neut_para_kkin0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_neut_para_cov002_kkin0.rds")
```

### kkpr 0 model

Here the same model is run, except the slope parameter, k, for VEpr is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEpr.

```{r kkpr0 model run}
trial_dat <- list(N=num_subjects_neuts, 
                  C=neut_cc_data$C,
                  A=neut_cc_data$A,
                  vaccinated=neut_cc_data$vaccinated,
                  Z=neut_cc_data$neuts.PB28_log10,
                  pers_yrs_at_risk=neut_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(neut_cc_data$neuts.PB28_log10),
                  max_Z=max(neut_cc_data$neuts.PB28_log10),
                  weight=neut_cc_data$neut_weight,
                  cov1=neut_cc_data$HCW_0,
                  cov2=neut_cc_data$HCW_1plus,
                  cov3=neut_cc_data$sc_age_cen,
                  cov4=neut_cc_data$obese,
                  cov5=neut_cc_data$white,
                  cov6=neut_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kkpr0.stan')
fit_neut_para_kkpr0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_neut_para_kkpr0)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_in", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_neut_para_kkpr0, pars=pars)
traceplot(fit_neut_para_kkpr0, pars=pars, inc_warmup = TRUE)
pairs(fit_neut_para_kkpr0, pars=pars)
```


```{r save as RDS}
fit_neut_para_kkpr0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_neut_para_kkpr0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_neut_para_cov002_kkpr0.rds")
```



### kkin and kkpr 0
Here the same model is run, except the slope parameter, k, for VEin and VEpr is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEin or VEpr.
```{r kk0 model run}
trial_dat <- list(N=num_subjects_neuts, 
                  C=neut_cc_data$C,
                  A=neut_cc_data$A,
                  vaccinated=neut_cc_data$vaccinated,
                  Z=neut_cc_data$neuts.PB28_log10,
                  pers_yrs_at_risk=neut_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(neut_cc_data$neuts.PB28_log10),
                  max_Z=max(neut_cc_data$neuts.PB28_log10),
                  weight=neut_cc_data$neut_weight,
                  cov1=neut_cc_data$HCW_0,
                  cov2=neut_cc_data$HCW_1plus,
                  cov3=neut_cc_data$sc_age_cen,
                  cov4=neut_cc_data$obese,
                  cov5=neut_cc_data$white,
                  cov6=neut_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kk0.stan')
fit_neut_para_kk0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_neut_para_kk0)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_neut_para_kk0, pars=pars)
traceplot(fit_neut_para_kk0, pars=pars, inc_warmup = TRUE)
pairs(fit_neut_para_kk0, pars=pars)
```


```{r save as RDS}
fit_neut_para_kk0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_neut_para_kk0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_neut_para_cov002_kk0.rds")
```



## b. Spike antibodies

The same models as above are run for the Spike binding antibodies outcome.

### Baseline model

```{r baseline model run}
#spike_cc_data$spike_rbd_weight<-1
trial_dat <- list(N=num_subjects_spike, 
                  C=spike_cc_data$C,
                  A=spike_cc_data$A,
                  vaccinated=spike_cc_data$vaccinated,
                  Z=spike_cc_data$igg.S.PB28_log10,
                  pers_yrs_at_risk=spike_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(spike_cc_data$igg.S.PB28_log10),
                  max_Z=max(spike_cc_data$igg.S.PB28_log10),
                  weight=spike_cc_data$spike_rbd_weight,
                  cov1=spike_cc_data$HCW_0,
                  cov2=spike_cc_data$HCW_1plus,
                  cov3=spike_cc_data$sc_age_cen,
                  cov4=spike_cc_data$obese,
                  cov5=spike_cc_data$white,
                  cov6=spike_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002.stan')
fit_spike_para <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_spike_para)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_in", "kk_pr", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")

print(fit_spike_para, pars=pars)
traceplot(fit_spike_para, pars=pars, inc_warmup = TRUE)
pairs(fit_spike_para, pars=pars)
```


```{r save as RDS}
fit_spike_para@stanmodel@dso <- new("cxxdso")
#saveRDS(fit_spike_para, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_spike_para_cov002.rds")
#saveRDS(fit_spike_para, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_spike_para_cov002_noweights.rds")
```


```{r posterior distributions}
est_params<-data.frame(rstan::extract(fit_spike_para, pars=pars))
hist(est_params$kk_in, br=100)
hist(est_params$kk_pr, br=100)
hist(est_params$kk_in, br=1000, xlim = c(0,3))
hist(est_params$kk_pr, br=1000, xlim = c(0,3))
hist(est_params$n_50_in, br=100)
hist(est_params$n_50_pr, br=100)
hist(est_params$VE_direct_in, br=100)
hist(est_params$VE_direct_pr, br=100)
hist(est_params$VE_indirect_in, br=100)
hist(est_params$VE_indirect_pr, br=100)
hist(est_params$ps, br=100)
hist(est_params$scale_in, br=100)
hist(est_params$scale_pr, br=100)
```

```{r summarise}
est_params_summary<-est_params %>%
  summarise(
    mean_ps=mean(ps),
    mean_foi_ref=mean(foi_ref),
    mean_foi_cov1=mean(foi_cov1),
    mean_foi_cov2=mean(foi_cov2),
    mean_foi_cov3=mean(foi_cov3),
    mean_foi_cov4=mean(foi_cov4),
    mean_foi_cov5=mean(foi_cov5),
    mean_foi_cov6=mean(foi_cov6),
    mean_kk_in=mean(kk_in),
    mean_kk_pr=mean(kk_pr),
    mean_n_50_in=mean(n_50_in),
    mean_n_50_pr=mean(n_50_pr),
    mean_scale_in=mean(scale_in),
    mean_scale_pr=mean(scale_pr),
    mean_VE_direct_in=mean(VE_direct_in),
    mean_VE_direct_pr=mean(VE_direct_pr),
    mean_VE_indirect_in=mean(VE_indirect_in_mean),
    mean_VE_indirect_pr=mean(VE_indirect_pr_mean),
    mean_p_direct_in=mean(p_direct_in),
    mean_p_direct_pr=mean(p_direct_pr),
    mean_p_indirect_in=1-mean(p_direct_in),
    mean_p_indirect_pr=1-mean(p_direct_pr),
    mean_VE_in=mean(VE_in_mean),
    mean_VE_pr=mean(VE_pr_mean),
    mean_VE_sym=mean(VE_sym_mean),
    mean_VE_asym=mean(VE_asym_mean),
    
    median_ps=median(ps),
    median_foi_ref=median(foi_ref),
    median_foi_cov1=median(foi_cov1),
    median_foi_cov2=median(foi_cov2),
    median_foi_cov3=median(foi_cov3),
    median_foi_cov4=median(foi_cov4),
    median_foi_cov5=median(foi_cov5),
    median_foi_cov6=median(foi_cov6), 
    median_kk_in=median(kk_in),
    median_kk_pr=median(kk_pr),
    median_n_50_in=median(n_50_in),
    median_n_50_pr=median(n_50_pr),
    median_scale_in=median(scale_in),
    median_scale_pr=median(scale_pr),
    median_VE_direct_in=median(VE_direct_in),
    median_VE_direct_pr=median(VE_direct_pr),
    median_VE_indirect_in=median(VE_indirect_in_mean),
    median_VE_indirect_pr=median(VE_indirect_pr_mean),
    median_p_direct_in=median(p_direct_in),
    median_p_direct_pr=median(p_direct_pr),
    median_VE_in=median(VE_in_mean),
    median_VE_pr=median(VE_pr_mean),
    median_VE_sym=median(VE_sym_mean),
    median_VE_asym=median(VE_asym_mean),
    
    lci_ps=quantile(ps, 0.025),
    lci_foi_ref=quantile(foi_ref, 0.025),
    lci_foi_cov1=quantile(foi_cov1, 0.025),
    lci_foi_cov2=quantile(foi_cov2, 0.025),
    lci_foi_cov3=quantile(foi_cov3, 0.025),
    lci_foi_cov4=quantile(foi_cov4, 0.025),
    lci_foi_cov5=quantile(foi_cov5, 0.025),
    lci_foi_cov6=quantile(foi_cov6, 0.025),
    lci_kk_in=quantile(kk_in, 0.025),
    lci_kk_pr=quantile(kk_pr, 0.025),
    lci_n_50_in=quantile(n_50_in, 0.025),
    lci_n_50_pr=quantile(n_50_pr, 0.025),
    lci_scale_in=quantile(scale_in, 0.025),
    lci_scale_pr=quantile(scale_pr, 0.025),
    lci_VE_direct_in=quantile(VE_direct_in, 0.025),
    lci_VE_direct_pr=quantile(VE_direct_pr, 0.025),
    lci_VE_indirect_in=quantile(VE_indirect_in_mean, 0.025),
    lci_VE_indirect_pr=quantile(VE_indirect_pr_mean, 0.025),
    lci_p_direct_in=quantile(p_direct_in, 0.025),
    lci_p_direct_pr=quantile(p_direct_pr, 0.025),
    lci_p_indirect_in=1-quantile(p_direct_in, 0.975),
    lci_p_indirect_pr=1-quantile(p_direct_pr, 0.975),
    lci_VE_in=quantile(VE_in_mean, 0.025),
    lci_VE_pr=quantile(VE_pr_mean, 0.025),
    lci_VE_sym=quantile(VE_sym_mean, 0.025),
    lci_VE_asym=quantile(VE_asym_mean, 0.025),
    
        
    uci_ps=quantile(ps, 0.975),
    uci_foi_ref=quantile(foi_ref, 0.975),
    uci_foi_cov1=quantile(foi_cov1, 0.975),
    uci_foi_cov2=quantile(foi_cov2, 0.975),
    uci_foi_cov3=quantile(foi_cov3, 0.975),
    uci_foi_cov4=quantile(foi_cov4, 0.975),
    uci_foi_cov5=quantile(foi_cov5, 0.975),
    uci_foi_cov6=quantile(foi_cov6, 0.975),
    uci_kk_in=quantile(kk_in, 0.975),
    uci_kk_pr=quantile(kk_pr, 0.975),
    uci_n_50_in=quantile(n_50_in, 0.975),
    uci_n_50_pr=quantile(n_50_pr, 0.975),
    uci_scale_in=quantile(scale_in, 0.975),
    uci_scale_pr=quantile(scale_pr, 0.975),
    uci_VE_direct_in=quantile(VE_direct_in, 0.975),
    uci_VE_direct_pr=quantile(VE_direct_pr, 0.975),
    uci_VE_indirect_in=quantile(VE_indirect_in_mean, 0.975),
    uci_VE_indirect_pr=quantile(VE_indirect_pr_mean, 0.975),
    uci_p_direct_in=quantile(p_direct_in, 0.975),
    uci_p_direct_pr=quantile(p_direct_pr, 0.975),
    uci_p_indirect_in=1-quantile(p_direct_in, 0.025),
    uci_p_indirect_pr=1-quantile(p_direct_pr, 0.025),
    uci_VE_in=quantile(VE_in_mean, 0.975),
    uci_VE_pr=quantile(VE_pr_mean, 0.975),
    uci_VE_sym=quantile(VE_sym_mean, 0.975),
    uci_VE_asym=quantile(VE_asym_mean, 0.975)
  )%>%
  round(digits=2)
est_params_summary
```


```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_in))+
  geom_histogram(binwidth = 0.02, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_in))+
  geom_histogram(binwidth = 0.2, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian()

ggplot(est_params, aes(x=kk_pr))+
  geom_histogram(binwidth = 0.02, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_pr))+
  geom_histogram(binwidth = 0.2, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian()
```

```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_in))+
  geom_density()+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_in))+
  geom_density()+
  theme_bw()+coord_cartesian()
```

```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_pr))+
  geom_density()+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_pr))+
  geom_density()+
  theme_bw()+coord_cartesian()
```


```{r read model}
fit_spike_para<-readRDS("fit_spike_para_cov002.rds")
```



### kkin 0 model
Here the same model is run, except the slope parameter, k, for VEin is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEin.

```{r kkin0 model run}
trial_dat <- list(N=num_subjects_RBD, 
                  C=spike_cc_data$C,
                  A=spike_cc_data$A,
                  vaccinated=spike_cc_data$vaccinated,
                  Z=spike_cc_data$igg.S.PB28_log10,
                  pers_yrs_at_risk=spike_cc_data$pers_yrs_at_risk_cop,
                  
                  min_Z=min(spike_cc_data$igg.S.PB28_log10),
                  max_Z=max(spike_cc_data$igg.S.PB28_log10),
                  weight=spike_cc_data$spike_rbd_weight,
                  cov1=spike_cc_data$HCW_0,
                  cov2=spike_cc_data$HCW_1plus,
                  cov3=spike_cc_data$sc_age_cen,
                  cov4=spike_cc_data$obese,
                  cov5=spike_cc_data$white,
                  cov6=spike_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kkin0.stan')

fit_spike_para_kkin0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_spike_para_kkin0)
loo <- loo(log_lik)
print(loo)
#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_pr", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_spike_para_kkin0, pars=pars)
traceplot(fit_spike_para_kkin0, pars=pars, inc_warmup = TRUE)
pairs(fit_spike_para_kkin0, pars=pars)
```


```{rsave as RDS}
fit_spike_para_kkin0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_spike_para_kkin0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_spike_para_cov002_kkin0.rds")
```

### kkpr 0 model
Here the same model is run, except the slope parameter, k, for VEpr is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEpr.

```{r kkpr0 model run}
trial_dat <- list(N=num_subjects_RBD, 
                  C=spike_cc_data$C,
                  A=spike_cc_data$A,
                  vaccinated=spike_cc_data$vaccinated,
                  Z=spike_cc_data$igg.S.PB28_log10,
                  pers_yrs_at_risk=spike_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(spike_cc_data$igg.S.PB28_log10),
                  max_Z=max(spike_cc_data$igg.S.PB28_log10),
                  weight=spike_cc_data$spike_rbd_weight,
                  cov1=spike_cc_data$HCW_0,
                  cov2=spike_cc_data$HCW_1plus,
                  cov3=spike_cc_data$sc_age_cen,
                  cov4=spike_cc_data$obese,
                  cov5=spike_cc_data$white,
                  cov6=spike_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kkpr0.stan')
fit_spike_para_kkpr0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_spike_para_kkpr0)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_in", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_spike_para_kkpr0, pars=pars)
traceplot(fit_spike_para_kkpr0, pars=pars, inc_warmup = TRUE)
pairs(fit_spike_para_kkpr0, pars=pars)
```

```{r save as RDS}
fit_spike_para_kkpr0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_spike_para_kkpr0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_spike_para_cov002_kkpr0.rds")
```


### kkin and kkpr 0
Here the same model is run, except the slope parameter, k, for VEin and VEpr is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEin or VEpr.

```{r kk0 model run}
trial_dat <- list(N=num_subjects_RBD, 
                  C=spike_cc_data$C,
                  A=spike_cc_data$A,
                  vaccinated=spike_cc_data$vaccinated,
                  Z=spike_cc_data$igg.S.PB28_log10,
                  pers_yrs_at_risk=spike_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(spike_cc_data$igg.S.PB28_log10),
                  max_Z=max(spike_cc_data$igg.S.PB28_log10),
                  weight=spike_cc_data$spike_rbd_weight,
                  cov1=spike_cc_data$HCW_0,
                  cov2=spike_cc_data$HCW_1plus,
                  cov3=spike_cc_data$sc_age_cen,
                  cov4=spike_cc_data$obese,
                  cov5=spike_cc_data$white,
                  cov6=spike_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kk0.stan')
fit_spike_para_kk0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_spike_para_kk0)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_spike_para_kk0, pars=pars)
traceplot(fit_spike_para_kk0, pars=pars, inc_warmup = TRUE)
pairs(fit_spike_para_kk0, pars=pars)
```


```{r save as RDS}
fit_spike_para_kk0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_spike_para_kk0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_spike_para_cov002_kk0.rds")
```


## c. RBD antibodies

The same models as above are run for the RBD binding antibodies outcome.

### Baseline model

```{r baseline model run}
RBD_cc_data$spike_rbd_weight<-1
trial_dat <- list(N=num_subjects_RBD, 
                  C=RBD_cc_data$C,
                  A=RBD_cc_data$A,
                  vaccinated=RBD_cc_data$vaccinated,
                  Z=RBD_cc_data$igg.RBD.PB28_log10,
                  pers_yrs_at_risk=RBD_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(RBD_cc_data$igg.RBD.PB28_log10),
                  max_Z=max(RBD_cc_data$igg.RBD.PB28_log10),
                  weight=RBD_cc_data$spike_rbd_weight,
                  cov1=RBD_cc_data$HCW_0,
                  cov2=RBD_cc_data$HCW_1plus,
                  cov3=RBD_cc_data$sc_age_cen,
                  cov4=RBD_cc_data$obese,
                  cov5=RBD_cc_data$white,
                  cov6=RBD_cc_data$comorbidity
)
setwd("C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates models")#desktop
getwd()

model <- stan_model(file = 'Parametric_model_cov002.stan')
fit_RBD_para <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_RBD_para)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_in", "kk_pr", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_RBD_para, pars=pars)
traceplot(fit_RBD_para, pars=pars, inc_warmup = TRUE)
pairs(fit_RBD_para, pars=pars)
```


```{r save as RDS}
fit_RBD_para@stanmodel@dso <- new("cxxdso")
#saveRDS(fit_RBD_para, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_RBD_para_cov002.rds")
#saveRDS(fit_RBD_para, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_RBD_para_cov002_noweights.rds")
```


```{r posterior distributions}
est_params<-data.frame(rstan::extract(fit_RBD_para, pars=pars))
hist(est_params$kk_in, br=100)
hist(est_params$kk_pr, br=100)
hist(est_params$kk_in, br=1000, xlim = c(0,3))
hist(est_params$kk_pr, br=1000, xlim = c(0,3))
hist(est_params$n_50_in, br=100)
hist(est_params$n_50_pr, br=100)
hist(est_params$VE_direct_in, br=100)
hist(est_params$VE_direct_pr, br=100)
hist(est_params$VE_indirect_in, br=100)
hist(est_params$VE_indirect_pr, br=100)
hist(est_params$ps, br=100)
hist(est_params$scale_in, br=100)
hist(est_params$scale_pr, br=100)
```

```{r summarise}
est_params_summary<-est_params %>%
  summarise(
    mean_ps=mean(ps),
    mean_foi_ref=mean(foi_ref),
    mean_foi_cov1=mean(foi_cov1),
    mean_foi_cov2=mean(foi_cov2),
    mean_foi_cov3=mean(foi_cov3),
    mean_foi_cov4=mean(foi_cov4),
    mean_foi_cov5=mean(foi_cov5),
    mean_foi_cov6=mean(foi_cov6),
    mean_kk_in=mean(kk_in),
    mean_kk_pr=mean(kk_pr),
    mean_n_50_in=mean(n_50_in),
    mean_n_50_pr=mean(n_50_pr),
    mean_scale_in=mean(scale_in),
    mean_scale_pr=mean(scale_pr),
    mean_VE_direct_in=mean(VE_direct_in),
    mean_VE_direct_pr=mean(VE_direct_pr),
    mean_VE_indirect_in=mean(VE_indirect_in_mean),
    mean_VE_indirect_pr=mean(VE_indirect_pr_mean),
    mean_p_direct_in=mean(p_direct_in),
    mean_p_direct_pr=mean(p_direct_pr),
    mean_p_indirect_in=1-mean(p_direct_in),
    mean_p_indirect_pr=1-mean(p_direct_pr),
    mean_VE_in=mean(VE_in_mean),
    mean_VE_pr=mean(VE_pr_mean),
    mean_VE_sym=mean(VE_sym_mean),
    mean_VE_asym=mean(VE_asym_mean),
    
    median_ps=median(ps),
    median_foi_ref=median(foi_ref),
    median_foi_cov1=median(foi_cov1),
    median_foi_cov2=median(foi_cov2),
    median_foi_cov3=median(foi_cov3),
    median_foi_cov4=median(foi_cov4),
    median_foi_cov5=median(foi_cov5),
    median_foi_cov6=median(foi_cov6), 
    median_kk_in=median(kk_in),
    median_kk_pr=median(kk_pr),
    median_n_50_in=median(n_50_in),
    median_n_50_pr=median(n_50_pr),
    median_scale_in=median(scale_in),
    median_scale_pr=median(scale_pr),
    median_VE_direct_in=median(VE_direct_in),
    median_VE_direct_pr=median(VE_direct_pr),
    median_VE_indirect_in=median(VE_indirect_in_mean),
    median_VE_indirect_pr=median(VE_indirect_pr_mean),
    median_p_direct_in=median(p_direct_in),
    median_p_direct_pr=median(p_direct_pr),
    median_VE_in=median(VE_in_mean),
    median_VE_pr=median(VE_pr_mean),
    median_VE_sym=median(VE_sym_mean),
    median_VE_asym=median(VE_asym_mean),
    
    lci_ps=quantile(ps, 0.025),
    lci_foi_ref=quantile(foi_ref, 0.025),
    lci_foi_cov1=quantile(foi_cov1, 0.025),
    lci_foi_cov2=quantile(foi_cov2, 0.025),
    lci_foi_cov3=quantile(foi_cov3, 0.025),
    lci_foi_cov4=quantile(foi_cov4, 0.025),
    lci_foi_cov5=quantile(foi_cov5, 0.025),
    lci_foi_cov6=quantile(foi_cov6, 0.025),
    lci_kk_in=quantile(kk_in, 0.025),
    lci_kk_pr=quantile(kk_pr, 0.025),
    lci_n_50_in=quantile(n_50_in, 0.025),
    lci_n_50_pr=quantile(n_50_pr, 0.025),
    lci_scale_in=quantile(scale_in, 0.025),
    lci_scale_pr=quantile(scale_pr, 0.025),
    lci_VE_direct_in=quantile(VE_direct_in, 0.025),
    lci_VE_direct_pr=quantile(VE_direct_pr, 0.025),
    lci_VE_indirect_in=quantile(VE_indirect_in_mean, 0.025),
    lci_VE_indirect_pr=quantile(VE_indirect_pr_mean, 0.025),
    lci_p_direct_in=quantile(p_direct_in, 0.025),
    lci_p_direct_pr=quantile(p_direct_pr, 0.025),
    lci_p_indirect_in=1-quantile(p_direct_in, 0.975),
    lci_p_indirect_pr=1-quantile(p_direct_pr, 0.975),
    lci_VE_in=quantile(VE_in_mean, 0.025),
    lci_VE_pr=quantile(VE_pr_mean, 0.025),
    lci_VE_sym=quantile(VE_sym_mean, 0.025),
    lci_VE_asym=quantile(VE_asym_mean, 0.025),
    
        
    uci_ps=quantile(ps, 0.975),
    uci_foi_ref=quantile(foi_ref, 0.975),
    uci_foi_cov1=quantile(foi_cov1, 0.975),
    uci_foi_cov2=quantile(foi_cov2, 0.975),
    uci_foi_cov3=quantile(foi_cov3, 0.975),
    uci_foi_cov4=quantile(foi_cov4, 0.975),
    uci_foi_cov5=quantile(foi_cov5, 0.975),
    uci_foi_cov6=quantile(foi_cov6, 0.975),
    uci_kk_in=quantile(kk_in, 0.975),
    ci_kk_pr=quantile(kk_pr, 0.975),
    uci_n_50_in=quantile(n_50_in, 0.975),
    uci_n_50_pr=quantile(n_50_pr, 0.975),
    uci_scale_in=quantile(scale_in, 0.975),
    uci_scale_pr=quantile(scale_pr, 0.975),
    uci_VE_direct_in=quantile(VE_direct_in, 0.975),
    uci_VE_direct_pr=quantile(VE_direct_pr, 0.975),
    uci_VE_indirect_in=quantile(VE_indirect_in_mean, 0.975),
    uci_VE_indirect_pr=quantile(VE_indirect_pr_mean, 0.975),
    uci_p_direct_in=quantile(p_direct_in, 0.975),
    uci_p_direct_pr=quantile(p_direct_pr, 0.975),
    uci_p_indirect_in=1-quantile(p_direct_in, 0.025),
    uci_p_indirect_pr=1-quantile(p_direct_pr, 0.025),
    uci_VE_in=quantile(VE_in_mean, 0.975),
    uci_VE_pr=quantile(VE_pr_mean, 0.975),
    uci_VE_sym=quantile(VE_sym_mean, 0.975),
    uci_VE_asym=quantile(VE_asym_mean, 0.975)
  )%>%
  round(digits=2)
est_params_summary
```


```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_in))+
  geom_histogram(binwidth = 0.02, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_in))+
  geom_histogram(binwidth = 0.2, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian()

ggplot(est_params, aes(x=kk_pr))+
  geom_histogram(binwidth = 0.02, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_pr))+
  geom_histogram(binwidth = 0.2, color = "black", fill="grey60")+
  theme_bw()+coord_cartesian()
```

```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_in))+
  geom_density()+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_in))+
  geom_density()+
  theme_bw()+coord_cartesian()
```

```{r visualise slope parameter k}
ggplot(est_params, aes(x=kk_pr))+
  geom_density()+
  theme_bw()+coord_cartesian(x=c(0,2))

ggplot(est_params, aes(x=kk_pr))+
  geom_density()+
  theme_bw()+coord_cartesian()
```


```{r read model}
fit_RBD_para<-readRDS("fit_RBD_para_cov002.rds")
```

```{r name fit}
fit<-fit_RBD_para
```

### kkin 0 model
Here the same model is run, except the slope parameter, k, for VEin is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEin.

```{r kkin0 model run}
trial_dat <- list(N=num_subjects_RBD, 
                  C=RBD_cc_data$C,
                  A=RBD_cc_data$A,
                  vaccinated=RBD_cc_data$vaccinated,
                  Z=RBD_cc_data$igg.RBD.PB28_log10,
                  pers_yrs_at_risk=RBD_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(RBD_cc_data$igg.RBD.PB28_log10),
                  max_Z=max(RBD_cc_data$igg.RBD.PB28_log10),
                  weight=RBD_cc_data$spike_rbd_weight,
                  cov1=RBD_cc_data$HCW_0,
                  cov2=RBD_cc_data$HCW_1plus,
                  cov3=RBD_cc_data$sc_age_cen,
                  cov4=RBD_cc_data$obese,
                  cov5=RBD_cc_data$white,
                  cov6=RBD_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kkin0.stan')

fit_RBD_para_kkin0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_RBD_para_kkin0)
loo <- loo(log_lik)
print(loo)
#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_pr", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_RBD_para_kkin0, pars=pars)
traceplot(fit_RBD_para_kkin0, pars=pars, inc_warmup = TRUE)
pairs(fit_RBD_para_kkin0, pars=pars)
```


```{r save as RDS}
fit_RBD_para_kkin0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_RBD_para_kkin0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_RBD_para_cov002_kkin0.rds")
```


### kkpr 0 model
Here the same model is run, except the slope parameter, k, for VEpr is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEpr.
```{r kkpr0 model run}
trial_dat <- list(N=num_subjects_RBD, 
                  C=RBD_cc_data$C,
                  A=RBD_cc_data$A,
                  vaccinated=RBD_cc_data$vaccinated,
                  Z=RBD_cc_data$igg.RBD.PB28_log10,
                  pers_yrs_at_risk=RBD_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(RBD_cc_data$igg.RBD.PB28_log10),
                  max_Z=max(RBD_cc_data$igg.RBD.PB28_log10),
                  weight=RBD_cc_data$spike_rbd_weight,
                  cov1=RBD_cc_data$HCW_0,
                  cov2=RBD_cc_data$HCW_1plus,
                  cov3=RBD_cc_data$sc_age_cen,
                  cov4=RBD_cc_data$obese,
                  cov5=RBD_cc_data$white,
                  cov6=RBD_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kkpr0.stan')
fit_RBD_para_kkpr0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_RBD_para_kkpr0)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_in", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_RBD_para_kkpr0, pars=pars)
traceplot(fit_RBD_para_kkpr0, pars=pars, inc_warmup = TRUE)
pairs(fit_RBD_para_kkpr0, pars=pars)
```

```{r save as RDS}
fit_RBD_para_kkpr0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_RBD_para_kkpr0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_RBD_para_cov002_kkpr.rds")
```


### kkin and kkpr 0

Here the same model is run, except the slope parameter, k, for VEin and VEpr is set to 0. It therefore assumes that there is no relationship between the candidate CoP and VEin or VEpr.
```{r kk0 model run}
trial_dat <- list(N=num_subjects_RBD, 
                  C=RBD_cc_data$C,
                  A=RBD_cc_data$A,
                  vaccinated=RBD_cc_data$vaccinated,
                  Z=RBD_cc_data$igg.RBD.PB28_log10,
                  pers_yrs_at_risk=RBD_cc_data$pers_yrs_at_risk_cop,
                  min_Z=min(RBD_cc_data$igg.RBD.PB28_log10),
                  max_Z=max(RBD_cc_data$igg.RBD.PB28_log10),
                  weight=RBD_cc_data$spike_rbd_weight,
                  cov1=RBD_cc_data$HCW_0,
                  cov2=RBD_cc_data$HCW_1plus,
                  cov3=RBD_cc_data$sc_age_cen,
                  cov4=RBD_cc_data$obese,
                  cov5=RBD_cc_data$white,
                  cov6=RBD_cc_data$comorbidity
)

model <- stan_model(file = 'Parametric_model_cov002_kk0.stan')
fit_RBD_para_kk0 <- sampling(object = model, data = trial_dat, chains = 4, cores = 4) 
```

```{r evaluate model}
#log likelihood
log_lik <- extract_log_lik(fit_RBD_para_kk0)
loo <- loo(log_lik)
print(loo)

#visualise parameters
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")
print(fit_RBD_para_kk0, pars=pars)
traceplot(fit_RBD_para_kk0, pars=pars, inc_warmup = TRUE)
pairs(fit_RBD_para_kk0, pars=pars)
```

```{r save as RDS}
fit_RBD_para_kk0@stanmodel@dso <- new("cxxdso")
saveRDS(fit_RBD_para_kk0, file = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/Model fits/fit_RBD_para_cov002_kk0.rds")
```

