---
title: "Plotting COV002 CoP results"
output: html_document
author: "Lucy Williams"
---

*Summary*
This file provides code for generating visualisations of the parametric and non-parametric models - VE estimates across CoP titres. It also provides code for estimating VE at given CoP values, notably the lower limit of quantification and the 95th percentile of the CoP distribution.

- The parametric plots are generated by taking the parameters estimated for each iteration and estimating the VE for a given measure of the CoP distribution. At each CoP measure, the distribution is summarised with the mean, median, and 95% confidence intervals.
- The non-parametric models are coded to generate estimates for VE at each CoP measure. Here we extract the mean and confidence intervals to plot.


```{r load packages}
library(cowplot)
library(tidyverse)
```


```{r read parametric models}
#to save time rerunning the models, save then load when needed

#main results
fit_neut_para<-readRDS("fit_neut_para_cov002.rds")
fit_spike_para<-readRDS("fit_spike_para_cov002.rds")
fit_RBD_para<-readRDS("fit_RBD_para_cov002.rds")

#sensitivity no weights
#fit_neut_para<-readRDS("fit_neut_para_cov002_noweights.rds")
#fit_spike_para<-readRDS("fit_spike_para_cov002_noweights.rds")
#fit_RBD_para<-readRDS("fit_RBD_para_cov002_noweights.rds")

```

```{r read nonparametric models}
#main results
fit_neut_nonpara<-readRDS("fit_neut_nonpara_cov002.rds")
fit_spike_nonpara<-readRDS("fit_spike_nonpara_cov002.rds")
fit_RBD_nonpara<-readRDS("fit_RBD_nonpara_cov002.rds")

#sensitivity no weights
#fit_neut_nonpara<-readRDS("fit_neut_nonpara_cov002_noweights.rds")
#fit_spike_nonpara<-readRDS("fit_spike_nonpara_cov002_noweights.rds")
#fit_RBD_nonpara<-readRDS("fit_RBD_nonpara_cov002_noweights.rds")
```

```{r antibody distribution}
neut_cc_data%>%
  subset(vaccinated == 1)%>%
  ggplot(., aes(neuts.PB28_log))+
  geom_histogram()+
  theme_bw()
neut_cc_data%>%
  ggplot(., aes(neuts.PB28_log))+
  geom_histogram()+
  theme_bw()

spike_cc_data%>%
  subset(vaccinated == 1)%>%
  ggplot(., aes(igg.S.PB28_log))+
  geom_histogram()+
  theme_bw()
spike_cc_data%>%
  ggplot(., aes(igg.S.PB28_log))+
  geom_histogram()+
  theme_bw()

RBD_cc_data%>%
  subset(vaccinated == 1)%>%
  ggplot(., aes(igg.RBD.PB28_log))+
  geom_histogram()+
  theme_bw()
RBD_cc_data%>%
  ggplot(., aes(igg.RBD.PB28_log))+
  geom_histogram()+
  theme_bw()
```


```{r lloq}
#Set lower limit of quantification for each antibody measurement
lloq_neut<-40
lloq_spike<-33
lloq_rbd<-204
```

```{r set x axis}
#log10 (vaccine arm)
neuts<-c(seq(min(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1]), round(max(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1]), 1)+0.0103,0.1)) #extended the sequence because it was cutting it slightly short 
spikes<-seq(min(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1]), round(max(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1])+0.147761, 1), 0.1)
RBDs<-seq(min(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1]), round(max(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1]), 1)+0.1086,0.1)

```


# A) PARAMETRIC

## 1. Neuts

Manually create histogram data
```{r Manually create neut histogram data}
breaks = seq(min(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)])-0.3, max(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)]+0.3), 0.1) #gave a wider range 
interval_counts <- as.data.frame(table(cut(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)], breaks = breaks)))
interval_counts$neuts= breaks[2:length(breaks)] #set the x axis to the lower bound
max(interval_counts$Freq)
max_count_10<- round(max(interval_counts$Freq), -1)
interval_counts$Freq_scaled=interval_counts$Freq/max_count_10
```

```{r create para neut ve data}
pars=c("ps", "foi_ref", "foi_cov1", "foi_cov2", "foi_cov3", "foi_cov4", "foi_cov5", "foi_cov6", "kk_in", "kk_pr", "n_50_in", "n_50_pr", "scale_in", "scale_pr", "VE_direct_in", "VE_direct_pr", "VE_indirect_in_mean", "VE_indirect_pr_mean", "p_direct_in", "p_direct_pr", "VE_in_mean", "VE_pr_mean", "VE_sym_mean", "VE_asym_mean")

est_params<-data.frame(rstan::extract(fit_neut_para, pars=pars))

neut_dat<-data.frame(matrix(NA, nrow = nrow(est_params)*length(neuts), ncol = 6))
colnames(neut_dat)<-c("neut", "iter", "ve_in", "ve_pr", "ve_sym", "ve_asym")
neut_dat$neut<-rep(neuts, nrow(est_params))
neut_dat$iter<-rep(seq(1:nrow(est_params)), each=length(neuts))


for(i in 1:nrow(neut_dat)){
  neut_dat$ve_in[i]<- est_params[neut_dat$iter[i], "p_direct_in"]*est_params[neut_dat$iter[i], "scale_in"] + ((1-est_params[neut_dat$iter[i], "p_direct_in"])*est_params[neut_dat$iter[i], "scale_in"])*(1/(1+exp(-est_params[neut_dat$iter[i], "kk_in"]*(neut_dat$neut[i]-est_params[neut_dat$iter[i], "n_50_in"]))))
  neut_dat$ve_pr[i]<- est_params[neut_dat$iter[i], "p_direct_pr"]*est_params[neut_dat$iter[i], "scale_pr"] + ((1-est_params[neut_dat$iter[i], "p_direct_pr"])*est_params[neut_dat$iter[i], "scale_pr"])*(1/(1+exp(-est_params[neut_dat$iter[i], "kk_pr"]*(neut_dat$neut[i]-est_params[neut_dat$iter[i], "n_50_pr"]))))

  neut_dat$ve_sym<- 1-(1-neut_dat$ve_in)*(1-neut_dat$ve_pr)
  neut_dat$ve_asym<- 1-((1-est_params[neut_dat$iter[i], "ps"])*(1-neut_dat$ve_in) + est_params[neut_dat$iter[i], "ps"]*(1-neut_dat$ve_in)*neut_dat$ve_pr) / (1-est_params[neut_dat$iter[i], "ps"])
  
}

neut_dat_plot<-neut_dat %>%
  group_by(neut)%>%
  summarise(
    mean_ve_in=mean(ve_in),
    mean_ve_pr=mean(ve_pr),
    mean_ve_sym=mean(ve_sym),
    mean_ve_asym=mean(ve_asym),
    median_ve_in=median(ve_in),
    median_ve_pr=median(ve_pr),
    median_ve_sym=median(ve_sym),
    median_ve_asym=median(ve_asym),
    
    lci_ve_in=quantile(ve_in, 0.025),
    lci_ve_pr=quantile(ve_pr, 0.025),
    lci_ve_sym=quantile(ve_sym, 0.025),
    lci_ve_asym=quantile(ve_asym, 0.025),
    
    uci_ve_in=quantile(ve_in, 0.975),
    uci_ve_pr=quantile(ve_pr, 0.975),
    uci_ve_sym=quantile(ve_sym, 0.975),
    uci_ve_asym=quantile(ve_asym, 0.975)
    
  )

```


```{r final para neut plot}
(pn1<-ggplot(neut_dat_plot)+
  #histogram
  geom_bar(data=interval_counts, aes(x=neuts, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=neuts, ymin=lci_ve_in, ymax=uci_ve_in),alpha=0.3, fill="darkblue", color = "darkblue", size=0.3) +
  geom_ribbon(aes(x=neuts, ymin=lci_ve_pr, ymax=uci_ve_pr),alpha=0.3, fill="#E69F00", color = "#E69F00", size=0.3) +
  geom_line(aes(neuts, mean_ve_in, colour="infection"), size=1) +
  geom_line(aes(neuts, mean_ve_pr, colour="progression"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,0.5))+
  scale_color_manual(name="", values = c("infection" = "darkblue", "progression" = "#E69F00")) +
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Neutralising Ab titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(neuts), max(neuts)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_neut), linetype="dashed"))

(pn2<-ggplot(neut_dat_plot)+
  #histogram
  geom_bar(data=interval_counts, aes(x=neuts, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=neuts, ymin=lci_ve_sym, ymax=uci_ve_sym),alpha=0.3, fill="darkgreen", color="darkgreen", size=0.3) +
  geom_ribbon(aes(x=neuts, ymin=lci_ve_asym, ymax=uci_ve_asym),alpha=0.3, fill="red", color="red", size=0.3) +
  geom_line(aes(neuts, mean_ve_sym, colour="symptomatic"), size=1) +
  geom_line(aes(neuts, mean_ve_asym, colour="asymptomatic"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,0.5))+
  scale_color_manual(name="", values = c("symptomatic" = "darkgreen", "asymptomatic" = "red") )+
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Neutralising Ab titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(neuts), max(neuts)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_neut), linetype="dashed"))

pn <- list(pn1, pn2)
(pn3<-plot_grid(plotlist=pn, ncol=2, align='h'))

```


```{r save or read plot dat}
getwd()
#write.csv(neut_dat_plot, "neut_dat_plot_cov002_oct_noweights_log10.csv", row.names = FALSE)#the code to create the parametric plot data takes a while to run, so worth saving a copy
neut_dat_plot<-read.csv("C:/Users/lucyr/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/neut_dat_plot.csv")
```

## 2. Spike

```{r Manually create spike histogram data}
breaks = seq(min(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)])-0.3, max(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)]+0.3), 0.1) #gave a wider range 
interval_counts <- as.data.frame(table(cut(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)], breaks = breaks)))
interval_counts$spikes= breaks[2:length(breaks)] #set the x axis to the lower bound
max(interval_counts$Freq)
max_count_10<- round(max(interval_counts$Freq), -1)
max_count_10<-150 #looked weird so using this instead
interval_counts$Freq_scaled=interval_counts$Freq/max_count_10
```

```{r create para spike ve data}
est_params<-data.frame(rstan::extract(fit_spike_para, pars=pars))
spike_dat<-data.frame(matrix(NA, nrow = nrow(est_params)*length(spikes), ncol = 6))
colnames(spike_dat)<-c("neut", "iter", "ve_in", "ve_pr", "ve_sym", "ve_asym")
spike_dat$neut<-rep(spikes, nrow(est_params))
spike_dat$iter<-rep(seq(1:nrow(est_params)), each=length(spikes))

for(i in 1:nrow(spike_dat)){
  spike_dat$ve_in[i]<- est_params[spike_dat$iter[i], "p_direct_in"]*est_params[spike_dat$iter[i], "scale_in"] + ((1-est_params[spike_dat$iter[i], "p_direct_in"])*est_params[spike_dat$iter[i], "scale_in"])*(1/(1+exp(-est_params[spike_dat$iter[i], "kk_in"]*(spike_dat$neut[i]-est_params[spike_dat$iter[i], "n_50_in"]))))
  spike_dat$ve_pr[i]<- est_params[spike_dat$iter[i], "p_direct_pr"]*est_params[spike_dat$iter[i], "scale_pr"] + ((1-est_params[spike_dat$iter[i], "p_direct_pr"])*est_params[spike_dat$iter[i], "scale_pr"])*(1/(1+exp(-est_params[spike_dat$iter[i], "kk_pr"]*(spike_dat$neut[i]-est_params[spike_dat$iter[i], "n_50_pr"]))))

  spike_dat$ve_sym<- 1-(1-spike_dat$ve_in)*(1-spike_dat$ve_pr)
  spike_dat$ve_asym<- 1-((1-est_params[spike_dat$iter[i], "ps"])*(1-spike_dat$ve_in) + est_params[spike_dat$iter[i], "ps"]*(1-spike_dat$ve_in)*spike_dat$ve_pr) / (1-est_params[spike_dat$iter[i], "ps"])
  
}

spike_dat_plot<-spike_dat %>%
  group_by(neut)%>%
  summarise(
    mean_ve_in=mean(ve_in),
    mean_ve_pr=mean(ve_pr),
    mean_ve_sym=mean(ve_sym),
    mean_ve_asym=mean(ve_asym),
    median_ve_in=median(ve_in),
    median_ve_pr=median(ve_pr),
    median_ve_sym=median(ve_sym),
    median_ve_asym=median(ve_asym),
    
    lci_ve_in=quantile(ve_in, 0.025),
    lci_ve_pr=quantile(ve_pr, 0.025),
    lci_ve_sym=quantile(ve_sym, 0.025),
    lci_ve_asym=quantile(ve_asym, 0.025),
    
    uci_ve_in=quantile(ve_in, 0.975),
    uci_ve_pr=quantile(ve_pr, 0.975),
    uci_ve_sym=quantile(ve_sym, 0.975),
    uci_ve_asym=quantile(ve_asym, 0.975)
    
  )

```

```{r final para spike plot}
(ps1<-ggplot(spike_dat_plot)+
  #histogram
  geom_bar(data=interval_counts, aes(x=spikes, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=spikes, ymin=lci_ve_in, ymax=uci_ve_in),alpha=0.3, fill="darkblue", color = "darkblue", size=0.3) +
  geom_ribbon(aes(x=spikes, ymin=lci_ve_pr, ymax=uci_ve_pr),alpha=0.3, fill="#E69F00", color = "#E69F00", size=0.3) +
  geom_line(aes(spikes, mean_ve_in, colour="infection"), size=1) +
  geom_line(aes(spikes, mean_ve_pr, colour="progression"), size=1) +
  scale_x_continuous(breaks=seq(1, 6,1))+
  scale_color_manual(name="", values = c("infection" = "darkblue", "progression" = "#E69F00")) +
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Spike IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(spikes), max(spikes)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_neut), linetype="dashed"))

(ps2<-ggplot(spike_dat_plot)+
  #histogram
  geom_bar(data=interval_counts, aes(x=spikes, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=spikes, ymin=lci_ve_sym, ymax=uci_ve_sym),alpha=0.3, fill="darkgreen", color="darkgreen", size=0.3) +
  geom_ribbon(aes(x=spikes, ymin=lci_ve_asym, ymax=uci_ve_asym),alpha=0.3, fill="red", color="red", size=0.3) +
  geom_line(aes(spikes, mean_ve_sym, colour="symptomatic"), size=1) +
  geom_line(aes(spikes, mean_ve_asym, colour="asymptomatic"), size=1) +
  scale_x_continuous(breaks=seq(1, 6,1))+
  scale_color_manual(name="", values = c("symptomatic" = "darkgreen", "asymptomatic" = "red") )+
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Spike IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(spikes), max(spikes)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_spike), linetype="dashed"))

ps <- list(ps1, ps2)
(ps3<-plot_grid(plotlist=ps, ncol=2, align='h'))

```

```{r save or read plot dat}
getwd()
#write.csv(spike_dat_plot, "spike_dat_plot_cov002_oct_log10.csv", row.names = FALSE)
spike_dat_plot<-read.csv("C:/Users/lucyr/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/spike_dat_plot.csv")
```

## 3. RBD
```{r Manually create RBD histogram data}
breaks = seq(min(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)])-0.3, max(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)]+0.3), 0.1) #gave a wider range 
interval_counts <- as.data.frame(table(cut(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)], breaks = breaks)))
interval_counts$RBDs= breaks[2:length(breaks)] #set the x axis to the lower bound
max(interval_counts$Freq)
max_count_10<- round(max(interval_counts$Freq), -1)
max_count_10<-150 #looked weird so using this instead
interval_counts$Freq_scaled=interval_counts$Freq/max_count_10
```

```{r create para RBD ve data}
est_params<-data.frame(rstan::extract(fit_RBD_para, pars=pars))
RBD_dat<-data.frame(matrix(NA, nrow = nrow(est_params)*length(RBDs), ncol = 6))
colnames(RBD_dat)<-c("neut", "iter", "ve_in", "ve_pr", "ve_sym", "ve_asym")
RBD_dat$neut<-rep(RBDs, nrow(est_params))
RBD_dat$iter<-rep(seq(1:nrow(est_params)), each=length(RBDs))



for(i in 1:nrow(RBD_dat)){
  RBD_dat$ve_in[i]<- est_params[RBD_dat$iter[i], "p_direct_in"]*est_params[RBD_dat$iter[i], "scale_in"] + ((1-est_params[RBD_dat$iter[i], "p_direct_in"])*est_params[RBD_dat$iter[i], "scale_in"])*(1/(1+exp(-est_params[RBD_dat$iter[i], "kk_in"]*(RBD_dat$neut[i]-est_params[RBD_dat$iter[i], "n_50_in"]))))
  RBD_dat$ve_pr[i]<- est_params[RBD_dat$iter[i], "p_direct_pr"]*est_params[RBD_dat$iter[i], "scale_pr"] + ((1-est_params[RBD_dat$iter[i], "p_direct_pr"])*est_params[RBD_dat$iter[i], "scale_pr"])*(1/(1+exp(-est_params[RBD_dat$iter[i], "kk_pr"]*(RBD_dat$neut[i]-est_params[RBD_dat$iter[i], "n_50_pr"]))))

  RBD_dat$ve_sym<- 1-(1-RBD_dat$ve_in)*(1-RBD_dat$ve_pr)
  RBD_dat$ve_asym<- 1-((1-est_params[RBD_dat$iter[i], "ps"])*(1-RBD_dat$ve_in) + est_params[RBD_dat$iter[i], "ps"]*(1-RBD_dat$ve_in)*RBD_dat$ve_pr) / (1-est_params[RBD_dat$iter[i], "ps"])
  
}

RBD_dat_plot<-RBD_dat %>%
  group_by(neut)%>%
  summarise(
    mean_ve_in=mean(ve_in),
    mean_ve_pr=mean(ve_pr),
    mean_ve_sym=mean(ve_sym),
    mean_ve_asym=mean(ve_asym),
    median_ve_in=median(ve_in),
    median_ve_pr=median(ve_pr),
    median_ve_sym=median(ve_sym),
    median_ve_asym=median(ve_asym),
    
    lci_ve_in=quantile(ve_in, 0.025),
    lci_ve_pr=quantile(ve_pr, 0.025),
    lci_ve_sym=quantile(ve_sym, 0.025),
    lci_ve_asym=quantile(ve_asym, 0.025),
    
    uci_ve_in=quantile(ve_in, 0.975),
    uci_ve_pr=quantile(ve_pr, 0.975),
    uci_ve_sym=quantile(ve_sym, 0.975),
    uci_ve_asym=quantile(ve_asym, 0.975)
    
  )

```

```{r final para RBD plot}
(pr1<-ggplot(RBD_dat_plot)+
  #histogram
  geom_bar(data=interval_counts, aes(x=RBDs, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=RBDs, ymin=lci_ve_in, ymax=uci_ve_in),alpha=0.3, fill="darkblue", color = "darkblue", size=0.3) +
  geom_ribbon(aes(x=RBDs, ymin=lci_ve_pr, ymax=uci_ve_pr),alpha=0.3, fill="#E69F00", color = "#E69F00", size=0.3) +
  geom_line(aes(RBDs, mean_ve_in, colour="infection"), size=1) +
  geom_line(aes(RBDs, mean_ve_pr, colour="progression"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,1))+
  scale_color_manual(name="", values = c("infection" = "darkblue", "progression" = "#E69F00")) +
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("RBD IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(RBDs), max(RBDs)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_rbd), linetype="dashed"))

(pr2<-ggplot(RBD_dat_plot)+
  #histogram
  geom_bar(data=interval_counts, aes(x=RBDs, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=RBDs, ymin=lci_ve_sym, ymax=uci_ve_sym),alpha=0.3, fill="darkgreen", color="darkgreen", size=0.3) +
  geom_ribbon(aes(x=RBDs, ymin=lci_ve_asym, ymax=uci_ve_asym),alpha=0.3, fill="red", color="red", size=0.3) +
  geom_line(aes(RBDs, mean_ve_sym, colour="symptomatic"), size=1) +
  geom_line(aes(RBDs, mean_ve_asym, colour="asymptomatic"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,1))+
  scale_color_manual(name="", values = c("symptomatic" = "darkgreen", "asymptomatic" = "red") )+
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("RBD IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(RBDs), max(RBDs)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_rbd), linetype="dashed"))

pr <- list(pr1, pr2)
(pr3<-plot_grid(plotlist=pr, ncol=2, align='h'))

```


```{r save or read plot dat}
getwd()
#write.csv(RBD_dat_plot, "RBD_dat_plot_cov002_oct_noweights.csv", row.names = FALSE)
RBD_dat_plot<-read.csv("C:/Users/lucyr/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/Correlates of protection/Correlates COV002/RBD_dat_plot.csv")
```


# B) NON-PARAMETRIC

## 1. Create data

```{r set x axis}
#log10 (vaccine arm)
neuts<-c(seq(min(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1]), round(max(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1]), 1)+0.0103,0.1)) #extended the sequence because it was cutting it slightly short 
spikes<-seq(min(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1]), round(max(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1])+0.147761, 1), 0.1)
RBDs<-seq(min(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1]), round(max(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1]), 1)+0.1086,0.1)
```

```{r set knots}
#log10
neut_knots=unname(quantile(neut_cc_data$neuts.PB28_log10[neut_cc_data$vaccinated==1],probs=seq(from=0, to=1, length.out = 3)))
spike_knots=unname(quantile(spike_cc_data$igg.S.PB28_log10[spike_cc_data$vaccinated==1],probs=seq(from=0, to=1, length.out = 3)))
RBD_knots=unname(quantile(RBD_cc_data$igg.RBD.PB28_log10[RBD_cc_data$vaccinated==1],probs=seq(from=0, to=1, length.out = 3)))
```

```{r data for plot neuts}
#VEin
cVE_in_hat<-rstan::extract(fit_neut_nonpara)[["cVE_hat"]]
cVE_in_hat_mean<-apply(cVE_in_hat,2,median)
cVE_in_hat_interval<-apply(cVE_in_hat,2,quantile,c(0.025,0.975))
#VEpr
cVE_pr_hat<-rstan::extract(fit_neut_nonpara)[["cVE_pr_hat"]]
cVE_pr_hat_mean<-apply(cVE_pr_hat,2,median)
cVE_pr_hat_interval<-apply(cVE_pr_hat,2,quantile,c(0.025,0.975))
#VEsym
cVE_sym_hat<-1-(1-cVE_in_hat)*(1-cVE_pr_hat)
cVE_sym_hat_mean<-apply(cVE_sym_hat,2,mean)
cVE_sym_hat_interval<-apply(cVE_sym_hat,2,quantile,c(0.025,0.975))
#VEasym
ps<-rstan::extract(fit_neut_nonpara)[["ps"]]
ps_mean<-mean(ps)
cVE_asym_hat<-   1-((1-ps_mean)*(1-cVE_in_hat) + ps_mean*(1-cVE_in_hat)*cVE_pr_hat) / (1-ps_mean)

cVE_asym_hat_mean<-apply(cVE_asym_hat,2,mean)
cVE_asym_hat_interval<-apply(cVE_asym_hat,2,quantile,c(0.025,0.975))

plot_dat_neut<-data.frame(cbind(neuts,cVE_in_hat_mean,t(cVE_in_hat_interval), cVE_pr_hat_mean,t(cVE_pr_hat_interval), cVE_sym_hat_mean,t(cVE_sym_hat_interval), cVE_asym_hat_mean,t(cVE_asym_hat_interval)))
names(plot_dat_neut)<-c("neuts","mean","lower","upper","mean_pr","lower_pr","upper_pr","mean_sym","lower_sym","upper_sym","mean_asym","lower_asym","upper_asym")
```


```{r data for plot spikes}
#VEin
cVE_in_hat<-rstan::extract(fit_spike_nonpara)[["cVE_hat"]]
cVE_in_hat_mean<-apply(cVE_in_hat,2,mean)
cVE_in_hat_interval<-apply(cVE_in_hat,2,quantile,c(0.025,0.975))
#VEpr
cVE_pr_hat<-rstan::extract(fit_spike_nonpara)[["cVE_pr_hat"]]
cVE_pr_hat_mean<-apply(cVE_pr_hat,2,mean)
cVE_pr_hat_interval<-apply(cVE_pr_hat,2,quantile,c(0.025,0.975))
#VEsym
cVE_sym_hat<-1-(1-cVE_in_hat)*(1-cVE_pr_hat)
cVE_sym_hat_mean<-apply(cVE_sym_hat,2,mean)
cVE_sym_hat_interval<-apply(cVE_sym_hat,2,quantile,c(0.025,0.975))
#VEasym
ps<-rstan::extract(fit_spike_nonpara)[["ps"]]
ps_mean<-mean(ps)
cVE_asym_hat<-   1-((1-ps_mean)*(1-cVE_in_hat) + ps_mean*(1-cVE_in_hat)*cVE_pr_hat) / (1-ps_mean)

cVE_asym_hat_mean<-apply(cVE_asym_hat,2,mean)
cVE_asym_hat_interval<-apply(cVE_asym_hat,2,quantile,c(0.025,0.975))

plot_dat_spike<-data.frame(cbind(spikes,cVE_in_hat_mean,t(cVE_in_hat_interval), cVE_pr_hat_mean,t(cVE_pr_hat_interval), cVE_sym_hat_mean,t(cVE_sym_hat_interval), cVE_asym_hat_mean,t(cVE_asym_hat_interval)))
names(plot_dat_spike)<-c("spikes","mean","lower","upper","mean_pr","lower_pr","upper_pr","mean_sym","lower_sym","upper_sym","mean_asym","lower_asym","upper_asym")
```

```{r data for plot RBDs}
#VEin
cVE_in_hat<-rstan::extract(fit_RBD_nonpara)[["cVE_hat"]]
cVE_in_hat_mean<-apply(cVE_in_hat,2,mean)
cVE_in_hat_interval<-apply(cVE_in_hat,2,quantile,c(0.025,0.975))
#VEpr
cVE_pr_hat<-rstan::extract(fit_RBD_nonpara)[["cVE_pr_hat"]]
cVE_pr_hat_mean<-apply(cVE_pr_hat,2,mean)
cVE_pr_hat_interval<-apply(cVE_pr_hat,2,quantile,c(0.025,0.975))
#VEsym
cVE_sym_hat<-1-(1-cVE_in_hat)*(1-cVE_pr_hat)
cVE_sym_hat_mean<-apply(cVE_sym_hat,2,mean)
cVE_sym_hat_interval<-apply(cVE_sym_hat,2,quantile,c(0.025,0.975))
#VEasym
ps<-rstan::extract(fit_RBD_nonpara)[["ps"]]
ps_mean<-mean(ps)
cVE_asym_hat<-   1-((1-ps_mean)*(1-cVE_in_hat) + ps_mean*(1-cVE_in_hat)*cVE_pr_hat) / (1-ps_mean)

cVE_asym_hat_mean<-apply(cVE_asym_hat,2,mean)
cVE_asym_hat_interval<-apply(cVE_asym_hat,2,quantile,c(0.025,0.975))

plot_dat_RBD<-data.frame(cbind(RBDs,cVE_in_hat_mean,t(cVE_in_hat_interval), cVE_pr_hat_mean,t(cVE_pr_hat_interval), cVE_sym_hat_mean,t(cVE_sym_hat_interval), cVE_asym_hat_mean,t(cVE_asym_hat_interval)))
names(plot_dat_RBD)<-c("RBDs","mean","lower","upper","mean_pr","lower_pr","upper_pr","mean_sym","lower_sym","upper_sym","mean_asym","lower_asym","upper_asym")
```


## 2. Neuts
Manually create histogram data
```{r Manually create neut histogram data}
breaks = seq(min(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)])-0.3, max(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)]+0.3), 0.1) #gave a wider range 
interval_counts <- as.data.frame(table(cut(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)], breaks = breaks)))
interval_counts$neuts= breaks[2:length(breaks)] #set the x axis to the lower bound
max(interval_counts$Freq)
max_count_10<- round(max(interval_counts$Freq), -1)
interval_counts$Freq_scaled=interval_counts$Freq/max_count_10
```

```{r final non para neut plot}
(pn1<-ggplot(plot_dat_neut)+
  #histogram
  geom_bar(data=interval_counts, aes(x=neuts, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=neuts, ymin=lower,ymax=upper),alpha=0.3, fill="darkblue", color = "darkblue", size=0.3) +
  geom_ribbon(aes(x=neuts, ymin=lower_pr,ymax=upper_pr),alpha=0.3, fill="#E69F00", color = "#E69F00", size=0.3) +
  geom_line(aes(neuts, mean, colour="infection"), size=1) +
  geom_line(aes(neuts, mean_pr, colour="progression"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,0.5))+
  scale_color_manual(name="", values = c("infection" = "darkblue", "progression" = "#E69F00")) +
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Neutralising Ab titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(neuts), max(neuts)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_neut), linetype="dashed"))

(pn2<-ggplot(plot_dat_neut)+
  #histogram
  geom_bar(data=interval_counts, aes(x=neuts, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=neuts, ymin=lower_sym,ymax=upper_sym),alpha=0.3, fill="darkgreen", color="darkgreen", size=0.3) +
  geom_ribbon(aes(x=neuts, ymin=lower_asym,ymax=upper_asym),alpha=0.3, fill="red", color="red", size=0.3) +
  geom_line(aes(neuts, mean_sym, colour="symptomatic"), size=1) +
  geom_line(aes(neuts, mean_asym, colour="asymptomatic"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,0.5))+
  scale_color_manual(name="", values = c("symptomatic" = "darkgreen", "asymptomatic" = "red") )+
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Neutralising Ab titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(neuts), max(neuts)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_neut), linetype="dashed"))

pn <- list(pn1, pn2)
(npn3<-plot_grid(plotlist=pn, ncol=2, align='h'))

```


## 3. Spike
```{r Manually create spike histogram data}
breaks = seq(min(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)])-0.3, max(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)]+0.3), 0.1) #gave a wider range 
interval_counts <- as.data.frame(table(cut(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)], breaks = breaks)))
interval_counts$spikes= breaks[2:length(breaks)] #set the x axis to the lower bound
max(interval_counts$Freq)
max_count_10<- round(max(interval_counts$Freq), -1)
max_count_10<-150 #looked weird so using this instead
interval_counts$Freq_scaled=interval_counts$Freq/max_count_10
```

```{r final non para spike plot}
(ps1<-ggplot(plot_dat_spike)+
  #histogram
  geom_bar(data=interval_counts, aes(x=spikes, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=spikes, ymin=lower,ymax=upper),alpha=0.3, fill="darkblue", color = "darkblue", size=0.3) +
  geom_ribbon(aes(x=spikes, ymin=lower_pr,ymax=upper_pr),alpha=0.3, fill="#E69F00", color = "#E69F00", size=0.3) +
  geom_line(aes(spikes, mean, colour="infection"), size=1) +
  geom_line(aes(spikes, mean_pr, colour="progression"), size=1) +
  scale_x_continuous(breaks=seq(1, 6,1))+
  scale_color_manual(name="", values = c("infection" = "darkblue", "progression" = "#E69F00")) +
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Spike IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(spikes), max(spikes)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_neut), linetype="dashed"))

(ps2<-ggplot(plot_dat_spike)+
  #histogram
  geom_bar(data=interval_counts, aes(x=spikes, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=spikes, ymin=lower_sym,ymax=upper_sym),alpha=0.3, fill="darkgreen", color="darkgreen", size=0.3) +
  geom_ribbon(aes(x=spikes, ymin=lower_asym,ymax=upper_asym),alpha=0.3, fill="red", color="red", size=0.3) +
  geom_line(aes(spikes, mean_sym, colour="symptomatic"), size=1) +
  geom_line(aes(spikes, mean_asym, colour="asymptomatic"), size=1) +
  scale_x_continuous(breaks=seq(1, 6,1))+
  scale_color_manual(name="", values = c("symptomatic" = "darkgreen", "asymptomatic" = "red") )+
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("Spike IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(spikes), max(spikes)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_spike), linetype="dashed"))

ps <- list(ps1, ps2)
(nps3<-plot_grid(plotlist=ps, ncol=2, align='h'))

```


## 4. RBD

```{r Manually create RBD histogram data}
breaks = seq(min(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)])-0.3, max(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)]+0.3), 0.1) #gave a wider range 
interval_counts <- as.data.frame(table(cut(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)], breaks = breaks)))
interval_counts$RBDs= breaks[2:length(breaks)] #set the x axis to the lower bound
max(interval_counts$Freq)
max_count_10<- round(max(interval_counts$Freq), -1)
max_count_10<-150 #looked weird so using this instead
interval_counts$Freq_scaled=interval_counts$Freq/max_count_10
```

```{r final non para RBD plot}
(pr1<-ggplot(plot_dat_RBD)+
  #histogram
  geom_bar(data=interval_counts, aes(x=RBDs, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=RBDs, ymin=lower,ymax=upper),alpha=0.3, fill="darkblue", color = "darkblue", size=0.3) +
  geom_ribbon(aes(x=RBDs, ymin=lower_pr,ymax=upper_pr),alpha=0.3, fill="#E69F00", color = "#E69F00", size=0.3) +
  geom_line(aes(RBDs, mean, colour="infection"), size=1) +
  geom_line(aes(RBDs, mean_pr, colour="progression"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,1))+
  scale_color_manual(name="", values = c("infection" = "darkblue", "progression" = "#E69F00")) +
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("RBD IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(RBDs), max(RBDs)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_rbd), linetype="dashed"))

(pr2<-ggplot(plot_dat_RBD)+
  #histogram
  geom_bar(data=interval_counts, aes(x=RBDs, y=Freq_scaled) ,fill="grey", alpha=0.5, stat = "identity", width=0.1, color="darkgrey")+
scale_y_continuous(
    sec.axis = sec_axis(~ . *max_count_10, name = "Count")
  )+
  #VE
  geom_ribbon(aes(x=RBDs, ymin=lower_sym,ymax=upper_sym),alpha=0.3, fill="darkgreen", color="darkgreen", size=0.3) +
  geom_ribbon(aes(x=RBDs, ymin=lower_asym,ymax=upper_asym),alpha=0.3, fill="red", color="red", size=0.3) +
  geom_line(aes(RBDs, mean_sym, colour="symptomatic"), size=1) +
  geom_line(aes(RBDs, mean_asym, colour="asymptomatic"), size=1) +
  scale_x_continuous(breaks=seq(-10,20,1))+
  scale_color_manual(name="", values = c("symptomatic" = "darkgreen", "asymptomatic" = "red") )+
  theme_minimal() +
  theme(legend.position="null", axis.title=element_blank(), text = element_text(size = 20), aspect.ratio=1) + xlab("RBD IgG titre (log)")+ #null/top for legend position
  coord_cartesian(x=c(min(RBDs), max(RBDs)+0.1), y=c(-1,1))+
  geom_vline(xintercept=log10(lloq_rbd), linetype="dashed"))

pr <- list(pr1, pr2)
(npr3<-plot_grid(plotlist=pr, ncol=2, align='h'))

```


# C) Generate VE estimates

Estimates for VE at given measures of the CoP (lower limit of quantification, 95th percentile)

## 1. Neuts

Visual approximations
```{r VE at a given titre neuts}
#this is approximating the VE at each titre from the data used to create the figure
summary(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)])#2.191 (mean log antibody titre in V arm)
log10(lloq_neut/2)#1.30103
log10(lloq_neut)# 1.60206
quantile(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)], 0.95) #95% of distribution above 2.93042, 99% above 3.276

#Parametric
neut_dat_plot#approximate
#mean log = 60.4% VEin and 37.3% VEpr
#lloq/2 = 57.6% VEin and 26.9% VEpr
#lloq = 58.4% VEin and 30.2 % VEpr
#max titre (3.4) 67.4% VEin and 48.7% VEpr
#95% of distribution (2.93) 63.7% VEin and 45.2% VEpr
#99% of distribution (3.276) 66.6% VEin and 48.05% VEpr

#Non-Parametric
plot_dat_neut
#mean log = 60.4% VEin and 38.0% VEpr
#lloq/2 = 64.7% VEin and 17.1% VEpr
#lloq = 56.8% VEin and  28.9% VEpr
#max titre (3.4) 65.2% VEin and 27.7% VEpr
#95% of distribution (2.93) 65.4% VEin and 39.0% VEpr
#99% of distribution (3.276) 65.5% VEin and 30.9% VEpr

#think would be best to go for the estimates at LLOQ and top 5% of distribution
lloq_neut_log10<-log10(lloq_neut)
neut_95_log10<-quantile(neut_cc_data$neuts.PB28_log10[which(neut_cc_data$vaccinated == 1)], 0.95)
```

Non-parametric estimation - assuming a linear relationship between the 0.1 intervals (used for manuscript)
```{r neut non para estimation lloq}
#LLOQ
plot_dat_neut[c(4,5),]
lloq_neut_log10
##VEin
#closest values to lloq_neut_log10 (1.60206)
multiplier<-((lloq_neut_log10-1.60103)/0.1) #1.03% of the way from the lower parts of the range (as a % of the difference between the top and bottom)

1.60103*1.0103
plot_dat_neut$neuts[c(4, 5)]
plot_dat_neut$mean[c(4, 5)]
plot_dat_neut$mean_pr[c(4, 5)]
difference_in<-(plot_dat_neut$mean[c(5)] - plot_dat_neut$mean[c(4)])*multiplier
difference_pr<-(plot_dat_neut$mean_pr[c(5)] - plot_dat_neut$mean_pr[c(4)])*multiplier
difference_in_lci<-(plot_dat_neut$lower[c(5)] - plot_dat_neut$lower[c(4)])*multiplier
difference_pr_lci<-(plot_dat_neut$lower_pr[c(5)] - plot_dat_neut$lower_pr[c(4)])*multiplier
difference_in_uci<-(plot_dat_neut$upper[c(5)] - plot_dat_neut$upper[c(4)])*multiplier
difference_pr_uci<-(plot_dat_neut$upper_pr[c(5)] - plot_dat_neut$upper_pr[c(4)])*multiplier

#VEin
round(plot_dat_neut$mean[c(4)]+difference_in, digits = 3)*100
round(plot_dat_neut$lower[c(4)]+difference_in_lci, digits = 3)*100
round(plot_dat_neut$upper[c(4)]+difference_in_uci, digits = 3)*100


#VEpr
round(plot_dat_neut$mean_pr[c(4)]+difference_pr, digits = 3)*100
round(plot_dat_neut$lower_pr[c(4)]+difference_pr_lci, digits = 3)*100
round(plot_dat_neut$upper_pr[c(4)]+difference_pr_uci, digits = 3)*100
```

```{r neut non para estimation 95 percentile}
#LLOQ
plot_dat_neut[c(17,18),]
neut_95_log10
##VEin
#closest values to neut_95_log10 (2.90103)
multiplier<-((neut_95_log10-2.90103)/0.1) #29% of the way from the lower parts of the range (as a % of the difference between the top and bottom)

plot_dat_neut$neuts[c(17,18)]
plot_dat_neut$mean[c(17,18)]
plot_dat_neut$mean_pr[c(17,18)]
difference_in<-(plot_dat_neut$mean[c(18)] - plot_dat_neut$mean[c(17)])*multiplier
difference_pr<-(plot_dat_neut$mean_pr[c(18)] - plot_dat_neut$mean_pr[c(17)])*multiplier
difference_in_lci<-(plot_dat_neut$lower[c(18)] - plot_dat_neut$lower[c(17)])*multiplier
difference_pr_lci<-(plot_dat_neut$lower_pr[c(18)] - plot_dat_neut$lower_pr[c(17)])*multiplier
difference_in_uci<-(plot_dat_neut$upper[c(18)] - plot_dat_neut$upper[c(17)])*multiplier
difference_pr_uci<-(plot_dat_neut$upper_pr[c(18)] - plot_dat_neut$upper_pr[c(17)])*multiplier

#VEin
round(plot_dat_neut$mean[c(17)]+difference_in, digits = 3)*100
round(plot_dat_neut$lower[c(17)]+difference_in_lci, digits = 3)*100
round(plot_dat_neut$upper[c(17)]+difference_in_uci, digits = 3)*100


#VEpr
round(plot_dat_neut$mean_pr[c(17)]+difference_pr, digits = 3)*100
round(plot_dat_neut$lower_pr[c(17)]+difference_pr_lci, digits = 3)*100
round(plot_dat_neut$upper_pr[c(17)]+difference_pr_uci, digits = 3)*100
```


Parametric estimation
```{r para neut ve estimation}
est_params<-data.frame(rstan::extract(fit_neut_para, pars=pars))
neut_dat<-data.frame(matrix(NA, nrow = nrow(est_params)*2, ncol = 6))#now generating 2 titre sets of estimates
colnames(neut_dat)<-c("neut", "iter", "ve_in", "ve_pr", "ve_sym", "ve_asym")
neut_dat$neut<-rep(c(lloq_neut_log10, neut_95_log10), nrow(est_params))
neut_dat$iter<-rep(seq(1:nrow(est_params)), each=2)



for(i in 1:nrow(neut_dat)){
  neut_dat$ve_in[i]<- est_params[neut_dat$iter[i], "p_direct_in"]*est_params[neut_dat$iter[i], "scale_in"] + ((1-est_params[neut_dat$iter[i], "p_direct_in"])*est_params[neut_dat$iter[i], "scale_in"])*(1/(1+exp(-est_params[neut_dat$iter[i], "kk_in"]*(neut_dat$neut[i]-est_params[neut_dat$iter[i], "n_50_in"]))))
  neut_dat$ve_pr[i]<- est_params[neut_dat$iter[i], "p_direct_pr"]*est_params[neut_dat$iter[i], "scale_pr"] + ((1-est_params[neut_dat$iter[i], "p_direct_pr"])*est_params[neut_dat$iter[i], "scale_pr"])*(1/(1+exp(-est_params[neut_dat$iter[i], "kk_pr"]*(neut_dat$neut[i]-est_params[neut_dat$iter[i], "n_50_pr"]))))

  neut_dat$ve_sym<- 1-(1-neut_dat$ve_in)*(1-neut_dat$ve_pr)
  neut_dat$ve_asym<- 1-((1-est_params[neut_dat$iter[i], "ps"])*(1-neut_dat$ve_in) + est_params[neut_dat$iter[i], "ps"]*(1-neut_dat$ve_in)*neut_dat$ve_pr) / (1-est_params[neut_dat$iter[i], "ps"])
  
}

neut_dat_ests<-neut_dat %>%
  group_by(neut)%>%
  summarise(
    mean_ve_in=mean(ve_in),
    lci_ve_in=quantile(ve_in, 0.025),
    uci_ve_in=quantile(ve_in, 0.975),
    mean_ve_pr=mean(ve_pr),
    lci_ve_pr=quantile(ve_pr, 0.025),
    uci_ve_pr=quantile(ve_pr, 0.975)
    
  )
#LLOQ mean and 95% CI VEin
round(neut_dat_ests$mean_ve_in[1]*100, digits =1)
round(neut_dat_ests$lci_ve_in[1]*100, digits =1)
round(neut_dat_ests$uci_ve_in[1]*100, digits =1)

#95% percentile VEin
round(neut_dat_ests$mean_ve_in[2]*100, digits =1)
round(neut_dat_ests$lci_ve_in[2]*100, digits =1)
round(neut_dat_ests$uci_ve_in[2]*100, digits =1)

#LLOQ mean and 95% CI VEpr
round(neut_dat_ests$mean_ve_pr[1]*100, digits =1)
round(neut_dat_ests$lci_ve_pr[1]*100, digits =1)
round(neut_dat_ests$uci_ve_pr[1]*100, digits =1)

#95% percentile VEpr
round(neut_dat_ests$mean_ve_pr[2]*100, digits =1)
round(neut_dat_ests$lci_ve_pr[2]*100, digits =1)
round(neut_dat_ests$uci_ve_pr[2]*100, digits =1)
```



## 2. Spikes
```{r VE at a given titre spikes}
#this is approximating the VE at each titre from the data used to create the figure
summary(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)])#4.489 (mean log antibody titre in V arm)
log10(lloq_spike/2)#1.217
log10(lloq_spike)# 1.5185
quantile(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)], 0.95) #95% of distribution above 5.08, 99% above 5.308 

#Parametric
spike_dat_plot#approximate
#lloq (1.5185) = 44.5% VEin and 30.0 % VEpr
#95% of distribution (5.08) 71.8% VEin and 44.2% VEpr

#Non-Parametric
plot_dat_spike
#lloq (1.5185) = 52.5% VEin and  30.0% VEpr
#95% of distribution (5.08) 72.8% VEin and 42.2% VEpr

lloq_spike_log10<-log10(lloq_spike)
spike_95_log10<-quantile(spike_cc_data$igg.S.PB28_log10[which(spike_cc_data$vaccinated == 1)], 0.95)

```

```{r spike non para estimation lloq}
#LLOQ
lloq_spike_log10
plot_dat_spike[c(4,5),]

##VEin
#closest values to lloq_spike_log10 (1.518514)
multiplier<-((lloq_spike_log10-1.517484	)/0.1) #1.03% of the way from the lower parts of the range (as a % of the difference between the top and bottom)

plot_dat_spike$spikes[c(4, 5)]
plot_dat_spike$mean[c(4, 5)]
plot_dat_spike$mean_pr[c(4, 5)]
difference_in<-(plot_dat_spike$mean[c(5)] - plot_dat_spike$mean[c(4)])*multiplier
difference_pr<-(plot_dat_spike$mean_pr[c(5)] - plot_dat_spike$mean_pr[c(4)])*multiplier
difference_in_lci<-(plot_dat_spike$lower[c(5)] - plot_dat_spike$lower[c(4)])*multiplier
difference_pr_lci<-(plot_dat_spike$lower_pr[c(5)] - plot_dat_spike$lower_pr[c(4)])*multiplier
difference_in_uci<-(plot_dat_spike$upper[c(5)] - plot_dat_spike$upper[c(4)])*multiplier
difference_pr_uci<-(plot_dat_spike$upper_pr[c(5)] - plot_dat_spike$upper_pr[c(4)])*multiplier

#VEin
round(plot_dat_spike$mean[c(4)]+difference_in, digits = 3)*100
round(plot_dat_spike$lower[c(4)]+difference_in_lci, digits = 3)*100
round(plot_dat_spike$upper[c(4)]+difference_in_uci, digits = 3)*100

#VEpr
round(plot_dat_spike$mean_pr[c(4)]+difference_pr, digits = 3)*100
round(plot_dat_spike$lower_pr[c(4)]+difference_pr_lci, digits = 3)*100
round(plot_dat_spike$upper_pr[c(4)]+difference_pr_uci, digits = 3)*100
```

```{r spike non para estimation 95 percentile}
#LLOQ
plot_dat_spike[c(39,40),]
spike_95_log10
##VEin
#closest values to spike_95_log10 (5.017484)
multiplier<-((spike_95_log10-5.017484)/0.1) #29% of the way from the lower parts of the range (as a % of the difference between the top and bottom)

plot_dat_spike$spikes[c(39,40)]
plot_dat_spike$mean[c(39,40)]
plot_dat_spike$mean_pr[c(39,40)]
difference_in<-(plot_dat_spike$mean[c(40)] - plot_dat_spike$mean[c(39)])*multiplier
difference_pr<-(plot_dat_spike$mean_pr[c(40)] - plot_dat_spike$mean_pr[c(39)])*multiplier
difference_in_lci<-(plot_dat_spike$lower[c(40)] - plot_dat_spike$lower[c(39)])*multiplier
difference_pr_lci<-(plot_dat_spike$lower_pr[c(40)] - plot_dat_spike$lower_pr[c(39)])*multiplier
difference_in_uci<-(plot_dat_spike$upper[c(40)] - plot_dat_spike$upper[c(39)])*multiplier
difference_pr_uci<-(plot_dat_spike$upper_pr[c(40)] - plot_dat_spike$upper_pr[c(39)])*multiplier

#VEin
round(plot_dat_spike$mean[c(39)]+difference_in, digits = 3)*100
round(plot_dat_spike$lower[c(39)]+difference_in_lci, digits = 3)*100
round(plot_dat_spike$upper[c(39)]+difference_in_uci, digits = 3)*100


#VEpr
round(plot_dat_spike$mean_pr[c(39)]+difference_pr, digits = 3)*100
round(plot_dat_spike$lower_pr[c(39)]+difference_pr_lci, digits = 3)*100
round(plot_dat_spike$upper_pr[c(39)]+difference_pr_uci, digits = 3)*100
```



```{r create para spike ve data}
est_params<-data.frame(rstan::extract(fit_spike_para, pars=pars))
spike_dat<-data.frame(matrix(NA, nrow = nrow(est_params)*2, ncol = 6))#now generating 2 titre sets of estimates
colnames(spike_dat)<-c("spike", "iter", "ve_in", "ve_pr", "ve_sym", "ve_asym")
spike_dat$spike<-rep(c(lloq_spike_log10, spike_95_log10), nrow(est_params))
spike_dat$iter<-rep(seq(1:nrow(est_params)), each=2)



for(i in 1:nrow(spike_dat)){
  spike_dat$ve_in[i]<- est_params[spike_dat$iter[i], "p_direct_in"]*est_params[spike_dat$iter[i], "scale_in"] + ((1-est_params[spike_dat$iter[i], "p_direct_in"])*est_params[spike_dat$iter[i], "scale_in"])*(1/(1+exp(-est_params[spike_dat$iter[i], "kk_in"]*(spike_dat$spike[i]-est_params[spike_dat$iter[i], "n_50_in"]))))
  spike_dat$ve_pr[i]<- est_params[spike_dat$iter[i], "p_direct_pr"]*est_params[spike_dat$iter[i], "scale_pr"] + ((1-est_params[spike_dat$iter[i], "p_direct_pr"])*est_params[spike_dat$iter[i], "scale_pr"])*(1/(1+exp(-est_params[spike_dat$iter[i], "kk_pr"]*(spike_dat$spike[i]-est_params[spike_dat$iter[i], "n_50_pr"]))))

  spike_dat$ve_sym<- 1-(1-spike_dat$ve_in)*(1-spike_dat$ve_pr)
  spike_dat$ve_asym<- 1-((1-est_params[spike_dat$iter[i], "ps"])*(1-spike_dat$ve_in) + est_params[spike_dat$iter[i], "ps"]*(1-spike_dat$ve_in)*spike_dat$ve_pr) / (1-est_params[spike_dat$iter[i], "ps"])
  
}

spike_dat_ests<-spike_dat %>%
  group_by(spike)%>%
  summarise(
    mean_ve_in=mean(ve_in),
    lci_ve_in=quantile(ve_in, 0.025),
    uci_ve_in=quantile(ve_in, 0.975),
    mean_ve_pr=mean(ve_pr),
    lci_ve_pr=quantile(ve_pr, 0.025),
    uci_ve_pr=quantile(ve_pr, 0.975)
    
  )
#LLOQ mean and 95% CI VEin
round(spike_dat_ests$mean_ve_in[1]*100, digits =1)
round(spike_dat_ests$lci_ve_in[1]*100, digits =1)
round(spike_dat_ests$uci_ve_in[1]*100, digits =1)

#95% percentile VEin
round(spike_dat_ests$mean_ve_in[2]*100, digits =1)
round(spike_dat_ests$lci_ve_in[2]*100, digits =1)
round(spike_dat_ests$uci_ve_in[2]*100, digits =1)

#LLOQ mean and 95% CI VEpr
round(spike_dat_ests$mean_ve_pr[1]*100, digits =1)
round(spike_dat_ests$lci_ve_pr[1]*100, digits =1)
round(spike_dat_ests$uci_ve_pr[1]*100, digits =1)

#95% percentile VEpr
round(spike_dat_ests$mean_ve_pr[2]*100, digits =1)
round(spike_dat_ests$lci_ve_pr[2]*100, digits =1)
round(spike_dat_ests$uci_ve_pr[2]*100, digits =1)
```


## 3. RBD
```{r VE at a given titre RBD}
#this is approximating the VE at each titre from the data used to create the figure
summary(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)])#4.615 (mean log antibody titre in V arm)
log10(lloq_rbd/2)#2.0086
log10(lloq_rbd)# 2.3096
quantile(RBD_cc_data$igg.RBD.PB28_log10[which(RBD_cc_data$vaccinated == 1)], 0.99) #95% of distribution above 5.25, 99% above 5.54 

#Parametric
RBD_dat_plot#approximate
#lloq (2.3096) = 46.9% VEin and 30.0 % VEpr
#95% of distribution (5.25) 71.8% VEin and 44.2% VEpr

#Non-Parametric
plot_dat_RBD
#lloq (2.3096) = 52.5% VEin and  30.0% VEpr
#95% of distribution (5.25) 71.9% VEin and 43.6% VEpr

lloq_RBD_log10<-log10(lloq_rbd)
RBD_95_log10<-quantile(RBD_cc_data$igg.S.PB28_log10[which(RBD_cc_data$vaccinated == 1)], 0.95)

```

```{r rbd non para estimation}
#LLOQ
lloq_RBD_log10
plot_dat_RBD[c(4,5),]

##VEin
#closest values to lloq_RBD_log10 (1.518514)
multiplier<-((lloq_RBD_log10-2.3086	)/0.1) #1.03% of the way from the lower parts of the range (as a % of the difference between the top and bottom)

plot_dat_RBD$RBDs[c(4, 5)]
plot_dat_RBD$mean[c(4, 5)]
plot_dat_RBD$mean_pr[c(4, 5)]
difference_in<-(plot_dat_RBD$mean[c(5)] - plot_dat_RBD$mean[c(4)])*multiplier
difference_pr<-(plot_dat_RBD$mean_pr[c(5)] - plot_dat_RBD$mean_pr[c(4)])*multiplier
difference_in_lci<-(plot_dat_RBD$lower[c(5)] - plot_dat_RBD$lower[c(4)])*multiplier
difference_pr_lci<-(plot_dat_RBD$lower_pr[c(5)] - plot_dat_RBD$lower_pr[c(4)])*multiplier
difference_in_uci<-(plot_dat_RBD$upper[c(5)] - plot_dat_RBD$upper[c(4)])*multiplier
difference_pr_uci<-(plot_dat_RBD$upper_pr[c(5)] - plot_dat_RBD$upper_pr[c(4)])*multiplier

#VEin
round(plot_dat_RBD$mean[c(4)]+difference_in, digits = 3)*100
round(plot_dat_RBD$lower[c(4)]+difference_in_lci, digits = 3)*100
round(plot_dat_RBD$upper[c(4)]+difference_in_uci, digits = 3)*100

#VEpr
round(plot_dat_RBD$mean_pr[c(4)]+difference_pr, digits = 3)*100
round(plot_dat_RBD$lower_pr[c(4)]+difference_pr_lci, digits = 3)*100
round(plot_dat_RBD$upper_pr[c(4)]+difference_pr_uci, digits = 3)*100
```

```{r RBD non para estimation 95 percentile}
#LLOQ
plot_dat_RBD[c(31,32),]
RBD_95_log10
##VEin
#closest values to RBD_95_log10 (5.080198 )
multiplier<-((RBD_95_log10-5.0086)/0.1) #29% of the way from the lower parts of the range (as a % of the difference between the top and bottom)

plot_dat_RBD$RBDs[c(31,32)]
plot_dat_RBD$mean[c(31,32)]
plot_dat_RBD$mean_pr[c(31,32)]
difference_in<-(plot_dat_RBD$mean[c(32)] - plot_dat_RBD$mean[c(31)])*multiplier
difference_pr<-(plot_dat_RBD$mean_pr[c(32)] - plot_dat_RBD$mean_pr[c(31)])*multiplier
difference_in_lci<-(plot_dat_RBD$lower[c(32)] - plot_dat_RBD$lower[c(31)])*multiplier
difference_pr_lci<-(plot_dat_RBD$lower_pr[c(32)] - plot_dat_RBD$lower_pr[c(31)])*multiplier
difference_in_uci<-(plot_dat_RBD$upper[c(32)] - plot_dat_RBD$upper[c(31)])*multiplier
difference_pr_uci<-(plot_dat_RBD$upper_pr[c(32)] - plot_dat_RBD$upper_pr[c(31)])*multiplier

#VEin
round(plot_dat_RBD$mean[c(31)]+difference_in, digits = 3)*100
round(plot_dat_RBD$lower[c(31)]+difference_in_lci, digits = 3)*100
round(plot_dat_RBD$upper[c(31)]+difference_in_uci, digits = 3)*100


#VEpr
round(plot_dat_RBD$mean_pr[c(32)]+difference_pr, digits = 3)*100
round(plot_dat_RBD$lower_pr[c(32)]+difference_pr_lci, digits = 3)*100
round(plot_dat_RBD$upper_pr[c(32)]+difference_pr_uci, digits = 3)*100
```


```{r create para RBD ve data}
est_params<-data.frame(rstan::extract(fit_RBD_para, pars=pars))
RBD_dat<-data.frame(matrix(NA, nrow = nrow(est_params)*2, ncol = 6))#now generating 2 titre sets of estimates
colnames(RBD_dat)<-c("neut", "iter", "ve_in", "ve_pr", "ve_sym", "ve_asym")
RBD_dat$RBD<-rep(c(lloq_RBD_log10, RBD_95_log10), nrow(est_params))
RBD_dat$iter<-rep(seq(1:nrow(est_params)), each=2)



for(i in 1:nrow(RBD_dat)){
  RBD_dat$ve_in[i]<- est_params[RBD_dat$iter[i], "p_direct_in"]*est_params[RBD_dat$iter[i], "scale_in"] + ((1-est_params[RBD_dat$iter[i], "p_direct_in"])*est_params[RBD_dat$iter[i], "scale_in"])*(1/(1+exp(-est_params[RBD_dat$iter[i], "kk_in"]*(RBD_dat$RBD[i]-est_params[RBD_dat$iter[i], "n_50_in"]))))
  RBD_dat$ve_pr[i]<- est_params[RBD_dat$iter[i], "p_direct_pr"]*est_params[RBD_dat$iter[i], "scale_pr"] + ((1-est_params[RBD_dat$iter[i], "p_direct_pr"])*est_params[RBD_dat$iter[i], "scale_pr"])*(1/(1+exp(-est_params[RBD_dat$iter[i], "kk_pr"]*(RBD_dat$RBD[i]-est_params[RBD_dat$iter[i], "n_50_pr"]))))

  RBD_dat$ve_sym<- 1-(1-RBD_dat$ve_in)*(1-RBD_dat$ve_pr)
  RBD_dat$ve_asym<- 1-((1-est_params[RBD_dat$iter[i], "ps"])*(1-RBD_dat$ve_in) + est_params[RBD_dat$iter[i], "ps"]*(1-RBD_dat$ve_in)*RBD_dat$ve_pr) / (1-est_params[RBD_dat$iter[i], "ps"])
  
}

RBD_dat_ests<-RBD_dat %>%
  group_by(RBD)%>%
  summarise(
    mean_ve_in=mean(ve_in),
    lci_ve_in=quantile(ve_in, 0.025),
    uci_ve_in=quantile(ve_in, 0.975),
    mean_ve_pr=mean(ve_pr),
    lci_ve_pr=quantile(ve_pr, 0.025),
    uci_ve_pr=quantile(ve_pr, 0.975)
    
  )
#LLOQ mean and 95% CI VEin
round(RBD_dat_ests$mean_ve_in[1]*100, digits =1)
round(RBD_dat_ests$lci_ve_in[1]*100, digits =1)
round(RBD_dat_ests$uci_ve_in[1]*100, digits =1)

#95% percentile VEin
round(RBD_dat_ests$mean_ve_in[2]*100, digits =1)
round(RBD_dat_ests$lci_ve_in[2]*100, digits =1)
round(RBD_dat_ests$uci_ve_in[2]*100, digits =1)

#LLOQ mean and 95% CI VEpr
round(RBD_dat_ests$mean_ve_pr[1]*100, digits =1)
round(RBD_dat_ests$lci_ve_pr[1]*100, digits =1)
round(RBD_dat_ests$uci_ve_pr[1]*100, digits =1)

#95% percentile VEpr
round(RBD_dat_ests$mean_ve_pr[2]*100, digits =1)
round(RBD_dat_ests$lci_ve_pr[2]*100, digits =1)
round(RBD_dat_ests$uci_ve_pr[2]*100, digits =1)
```



